---
title: ""
message: false
warning: false
---

# Packages

```{r}

library(gsheet) # Importação dos dados na nuvem
library(dplyr) # Manipulação 
library(readxl) # Improtação de dados no computador
#library(vegan) # Gráfico de HeatMap
library(ggthemes)  #Tema de fundo do gráfico
library(tidyverse) # Conjunto de pacotes
library(cowplot) # theme
library(writexl) #Exportação de dados
library(ggalluvial)
library(ggplot2) # Gráficos
library(lme4) # GLM
library(glmmTMB) #GLMM
library(rnaturalearth)
library(ggplot2)
library(ggridges)
library(performance)
library(DHARMa)
library(PresenceAbsence)
library(caret)
library(lmtest)
```

# Incidence - Logistic regression

```{r}
incidence = gsheet2tbl("https://docs.google.com/spreadsheets/d/1gr-TUpB-mNAhZVOBEcmHRwE1etorOeJvlhjcu7wEpz4/edit?usp=sharing")
```


```{r}

incidence =  incidence %>% 
 filter(soil %in% c("Argissolo","Cambissolo","Latossolo"))

#incidence = incidence %>% 
 # filter(!city %in% c("DFL","DFC"))

incidence[incidence$soil== "Argissolo",c("soil_corrigido")] <- 2
incidence[incidence$soil== "Cambissolo",c("soil_corrigido")] <- 5
incidence[incidence$soil== "Latossolo",c("soil_corrigido")] <- 6

incidence[incidence$cropping== "Corn_soybean",c("cropping_corrigido")] <- 1
incidence[incidence$cropping== "Corn_others",c("cropping_corrigido")] <- 2

incidence[incidence$tillage== "No-till",c("tillage_corrigido")] <- 1
incidence[incidence$tillage== "Minimum",c("tillage_corrigido")] <- 1
incidence[incidence$tillage== "Conventional",c("tillage_corrigido")] <- 2

incidence[incidence$region== "South",c("region_corrigido")] <- 1
incidence[incidence$region== "Southwest",c("region_corrigido")] <- 2
incidence[incidence$region== "East",c("region_corrigido")] <- 3
incidence[incidence$region== "Center",c("region_corrigido")] <- 4



```


# Selecting of variables (LASSO)

```{r}
library(glmnet)

all_genus = incidence |> 
   dplyr::select(incidence,sand,silt,clay,ec,ph,cropping_corrigido,tillage_corrigido,altitute,soil_corrigido)


set.seed(123)
lambdas <- 10^seq(2, -3, by = -.1)

y <- all_genus %>%
  dplyr::select(incidence) %>%
  as.matrix()



X <- all_genus%>%
  dplyr::select(-incidence) %>%
  as.matrix()


# Setting alpha = 1 implements lasso regression
lasso_reg_1week <- cv.glmnet(X, y,
  alpha = 1,
  family = "binomial",
  lambda = lambdas,
  standardize = TRUE,
  nfolds = 5
)
plot(lasso_reg_1week)

# Best
lambda_best_1week <- lasso_reg_1week$lambda.min
lambda_best_1week

lasso_model_1week <- glmnet(X,
  y,
  alpha = 1,
  family = "binomial",
  lambda = lambda_best_1week,
  standardize = TRUE
)
coef(lasso_model_1week)
assess.glmnet(lasso_model_1week,
  newx = X,
  newy = y
)

plot(roc.glmnet(lasso_model_1week,
  newx = X,
  newy = factor(y)
),
type = "l"
) # produces
```


# Framework modeling

```{r}

pratylenchus_log = incidence %>% 
  filter(genus == "Pratylenchus") %>% 
    filter(soil_corrigido %in% c("2","5","6")) %>% 
   filter(region_corrigido %in% c("1","2","3","4")) 
  


praty_log = glm(incidence ~  altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = pratylenchus_log, 
                family = binomial(link = "logit"))

check_singularity(praty_log)
model_performance(praty_log)
check_model(praty_log)


#with(summary(praty_log), 1 - deviance/null.deviance)


#----------------------------------------------------------

helicotylenchus_log = incidence %>% 
  filter(genus == "Helicotylenchus")%>% 
    filter(soil_corrigido %in% c("2","5","6"))%>% 
   filter(region_corrigido %in% c("1","2","3","4"))

heli_log = glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph,  data = helicotylenchus_log, 
                family = binomial(link = "logit"))

check_model(heli_log)
model_performance(heli_log)


#----------------------------------------------------------

rotylenchulus_log = incidence %>% 
  filter(genus == "Rotylenchulus")%>% 
   filter(soil_corrigido %in% c("2","5","6"))%>% 
   filter(region_corrigido %in% c("1","2","3","4"))

is.na(rotylenchulus_log)

rotyl_log = glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = rotylenchulus_log, 
                family = binomial(link = "logit"))

check_model(rotyl_log)
model_performance(rotyl_log)


#----------------------------------------------------------

criconematidae_log = incidence %>% 
  filter(genus == "Criconematidae")%>% 
    filter(soil_corrigido %in% c("2","5","6"))%>% 
   filter(region_corrigido %in% c("1","2","3","4"))

crico_log = glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = criconematidae_log, 
                family = binomial(link = "logit"))

check_model(rotyl_log)
model_performance(rotyl_log)


#----------------------------------------------------------

aphelenchus_log = incidence %>% 
  filter(genus == "Aphelenchus")%>% 
    filter(soil_corrigido %in% c("2","5","6"))%>% 
   filter(region_corrigido %in% c("1","2","3","4"))

aphel_log = glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchus_log, 
                family = binomial(link = "logit"))

check_model(aphel_log)
model_performance(aphel_log)


#----------------------------------------------------------

aphelenchoides_log = incidence %>% 
  filter(genus == "Aphelenchoides")%>% 
    filter(soil_corrigido %in% c("2","5","6"))%>% 
   filter(region_corrigido %in% c("1","2","3","4"))

aphels_log = glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchoides_log, 
                family = binomial(link = "logit"))

check_model(aphels_log)
model_performance(aphels_log)

```

## Wald x² and LRT - Adicionar linha para cada terms = ...

```{r}

#Pratylenchus

DAAG::vif(praty_log)
coef(praty_log)
summary(praty_log)

aod::wald.test(Sigma = vcov(praty_log), b = coef(praty_log), Terms =8)
lrtest(praty_log)


#Helicotylenchus

summary(heli_log)
aod::wald.test(Sigma = vcov(heli_log), b = coef(heli_log), Terms = 8)
lrtest (heli_log)


#Rotylenchulus

summary(rotyl_log)
aod::wald.test(Sigma = vcov(rotyl_log), b = coef(rotyl_log), Terms = 8)
lrtest (rotyl_log)


#Criconematidae

summary(crico_log)
coef(crico_log)
aod::wald.test(Sigma = vcov(crico_log), b = coef(crico_log), Terms = 8)
lrtest(crico_log)


#Aphelenchus

DAAG::vif(aphel_log)
summary(aphel_log)
coef(aphel_log)
aod::wald.test(Sigma = vcov(aphel_log), b = coef(aphel_log), Terms = 8)
lrtest (aphel_log)


#Aphelenchoides

summary(aphels_log)
coef(aphels_log)
aod::wald.test(Sigma = vcov(aphels_log), b = coef(aphels_log), Terms = 8)
lrtest (aphels_log)


```

## Models diagnostic (graphic)

```{r}
check_singularity(praty_log)
check_model(praty_log)

simulationOutput <- simulateResiduals(fittedModel = praty_log)
plot(simulationOutput)
```

### Aphelenchus

```{r}

#Model log with best
aphelenchus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchus_log,family = binomial(link = "logit"))

summary(aphelenchus_binomial)

pred_aphelenchus <- predict(aphelenchus_binomial, aphelenchus_log, type = "response")


#Better thresholds
dat_aphelenchus <- data.frame(1, aphelenchus_log$incidence, pred_aphelenchus)
  optimal.thresholds(dat_aphelenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_aphelenchus > 0.82)),  mode= "everything",  reference = as.factor(aphelenchus_log$incidence))


#Plot
auc.roc.plot(dat_aphelenchus)
calibration.plot(dat_aphelenchus)

error.threshold.plot(dat_aphelenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_aphelenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_aphelenchus,threshold=0.82,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_aphelenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_aphelenchus <- roc.plot.calculate(dat_aphelenchus) #data frame

roc_aphelenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_aphelenchus <- NULL
for(i in 1:nrow(aphelenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- aphelenchus_log[-i,]
    test <- aphelenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.82,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_aphelenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_aphelenchus)


```

### Aphelenchoides

```{r}
#Model log with best
aphelenchoides_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchoides_log,family = binomial(link = "logit"))

summary(aphelenchoides_binomial)

pred_aphelenchoides <- predict(aphelenchoides_binomial, aphelenchoides_log, type = "response")

#Better thresholds
dat_aphelenchoides <- data.frame(1, aphelenchoides_log$incidence, pred_aphelenchoides)
optimal.thresholds(dat_aphelenchoides)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_aphelenchoides > 0.65)),  mode= "everything",  reference = as.factor(aphelenchoides_log$incidence))


#Plot
auc.roc.plot(dat_aphelenchoides)
calibration.plot(dat_aphelenchoides)

error.threshold.plot(dat_aphelenchoides, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_aphelenchoides), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_aphelenchoides,threshold=0.65,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_aphelenchoides) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_aphelenchoides <- roc.plot.calculate(dat_aphelenchoides) #data frame

roc_aphelenchoides |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_aphelenchoides <- NULL
for(i in 1:nrow(aphelenchoides_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- aphelenchoides_log[-i,]
    test <- aphelenchoides_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchoides_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.65,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_aphelenchoides[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_aphelenchoides)


```

### Criconematidae

```{r}
#Model log with best
criconematidae_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = criconematidae_log,family = binomial(link = "logit"))

summary(criconematidae_binomial)

pred_criconematidae <- predict(criconematidae_binomial, criconematidae_log, type = "response")

#Better thresholds
dat_criconematidae <- data.frame(1, criconematidae_log$incidence, pred_criconematidae)
optimal.thresholds(dat_criconematidae)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_criconematidae > 0.26)),  mode= "everything",  reference = as.factor(criconematidae_log$incidence))


#Plot
auc.roc.plot(dat_criconematidae)
calibration.plot(dat_criconematidae)

error.threshold.plot(dat_criconematidae, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_criconematidae), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_criconematidae,threshold=0.26,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_criconematidae) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_criconematidae <- roc.plot.calculate(dat_criconematidae) #data frame

roc_criconematidae |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_criconematidae <- NULL
for(i in 1:nrow(criconematidae_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- criconematidae_log[-i,]
    test <- criconematidae_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = criconematidae_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.26,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_criconematidae[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_criconematidae)


```

### Helicotylenchus

```{r}
#Model log with best
helicotylenchus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = helicotylenchus_log,family = binomial(link = "logit"))

summary(helicotylenchus_binomial)

pred_helicotylenchus <- predict(helicotylenchus_binomial, helicotylenchus_log, type = "response")

#Better thresholds
dat_helicotylenchus <- data.frame(1, helicotylenchus_log$incidence, pred_helicotylenchus)
optimal.thresholds(dat_helicotylenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_helicotylenchus > 0.9)),  mode= "everything",  reference = as.factor(helicotylenchus_log$incidence))


#Plot
auc.roc.plot(dat_helicotylenchus)
calibration.plot(dat_helicotylenchus)

error.threshold.plot(dat_helicotylenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_helicotylenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_helicotylenchus,threshold=.9,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_helicotylenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_helicotylenchus <- roc.plot.calculate(dat_helicotylenchus) #data frame

roc_helicotylenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_helicotylenchus <- NULL
for(i in 1:nrow(helicotylenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- helicotylenchus_log[-i,]
    test <- helicotylenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = helicotylenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.9,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_helicotylenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_helicotylenchus)



```

### Rotylenchulus

```{r}
#Model log with best
rotylenchulus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = rotylenchulus_log,family = binomial(link = "logit"))

summary(rotylenchulus_binomial)

pred_rotylenchulus <- predict(rotylenchulus_binomial, rotylenchulus_log, type = "response")

#Better thresholds
dat_rotylenchulus <- data.frame(1, rotylenchulus_log$incidence, pred_rotylenchulus)
optimal.thresholds(dat_rotylenchulus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_rotylenchulus > .06)),  mode= "everything",  reference = as.factor(rotylenchulus_log$incidence))


#Plot
auc.roc.plot(dat_rotylenchulus)
calibration.plot(dat_rotylenchulus)

error.threshold.plot(dat_rotylenchulus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_rotylenchulus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_rotylenchulus,threshold=.06,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_rotylenchulus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_rotylenchulus <- roc.plot.calculate(dat_rotylenchulus) #data frame

roc_rotylenchulus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_rotylenchulus <- NULL
for(i in 1:nrow(rotylenchulus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- rotylenchulus_log[-i,]
    test <- rotylenchulus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = rotylenchulus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > .06,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_rotylenchulus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_rotylenchulus)

```

### Pratylenchus

```{r}
#Model log with best
pratylenchus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = pratylenchus_log,family = binomial(link = "logit"))

summary(pratylenchus_binomial)

pred_pratylenchus <- predict(pratylenchus_binomial, pratylenchus_log, type = "response")

#Better thresholds
dat_pratylenchus <- data.frame(1, pratylenchus_log$incidence, pred_pratylenchus)
optimal.thresholds(dat_pratylenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_pratylenchus > 0.58)),  mode= "everything",  reference = as.factor(pratylenchus_log$incidence))


#Plot
auc.roc.plot(dat_pratylenchus)
calibration.plot(dat_pratylenchus)

error.threshold.plot(dat_pratylenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_pratylenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_pratylenchus,threshold=.58,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_pratylenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_pratylenchus <- roc.plot.calculate(dat_pratylenchus) #data frame

roc_pratylenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_pratylenchus <- NULL
for(i in 1:nrow(pratylenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- pratylenchus_log[-i,]
    test <- pratylenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = pratylenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.58,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_pratylenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_pratylenchus)



```

## Joining

```{r}

loocv_all = cbind(acc_aphelenchoides,acc_aphelenchus,acc_criconematidae,
                  acc_helicotylenchus,acc_pratylenchus,acc_rotylenchulus)
```


## Odds ratio

### By cropping

```{r}


#Helicotylenchus

helicotylenchus_log$predicted = predict(heli_log, type = "response")
#helicotylenchus_log$predicted = exp(helicotylenchus_log$predicted)
helicotylenchus_log$predicted = (helicotylenchus_log$predicted/(1 - helicotylenchus_log$predicted))

helicotylenchus_log01_cropping<-helicotylenchus_log$predicted[helicotylenchus_log$cropping_corrigido==1]
helicotylenchus_log02_cropping<-helicotylenchus_log$predicted[helicotylenchus_log$cropping_corrigido==2]

helicotylenchus_log01_cropping = mean(helicotylenchus_log01_cropping)
helicotylenchus_log02_cropping = mean(helicotylenchus_log02_cropping)


odd_helicotylenchus_cropping = c(helicotylenchus_log01_cropping,helicotylenchus_log02_cropping)


#logodd_diff<-s1/s2


odd_helicotylenchus_cropping = as.data.frame(odd_helicotylenchus_cropping)
odd_helicotylenchus_cropping$nematode = c("Helicotylenchus","Helicotylenchus")
colnames(odd_helicotylenchus_cropping) = c("odd","nematode")
#odd_helicotylenchus_cropping$odd<-exp(odd_helicotylenchus_cropping$odd)

#----------------------------------------------------------

#Aphelenchus

aphelenchus_log$predicted = predict(aphel_log, type = "response")
#aphelenchus_log$predicted = exp(aphelenchus_log$predicted)
aphelenchus_log$predicted = (aphelenchus_log$predicted/(1 - aphelenchus_log$predicted))

aphelenchus_log01_cropping<-aphelenchus_log$predicted[aphelenchus_log$cropping_corrigido==1]
aphelenchus_log02_cropping<-aphelenchus_log$predicted[aphelenchus_log$cropping_corrigido==2]

aphelenchus_log01_cropping = mean(aphelenchus_log01_cropping)
aphelenchus_log02_cropping = mean(aphelenchus_log02_cropping)

odd_aphelenchus_cropping = c(aphelenchus_log01_cropping,aphelenchus_log02_cropping)


odd_aphelenchus_cropping = as.data.frame(odd_aphelenchus_cropping)
odd_aphelenchus_cropping$nematode = c("Aphelenchus","Aphelenchus")
colnames(odd_aphelenchus_cropping) = c("odd","nematode")
#odd_aphelenchus_cropping$odd<-exp(odd_aphelenchus_cropping$odd)

#----------------------------------------------------------

#Aphelenchoides

aphelenchoides_log$predicted = predict(aphels_log, type = "response")
#aphelenchoides_log$predicted = exp(aphelenchoides_log$predicted)
aphelenchoides_log$predicted = (aphelenchoides_log$predicted/(1 - aphelenchoides_log$predicted))

aphelenchoides_log01_cropping<-aphelenchoides_log$predicted[aphelenchoides_log$cropping_corrigido==1]
aphelenchoides_log02_cropping<-aphelenchoides_log$predicted[aphelenchoides_log$cropping_corrigido==2]

aphelenchoides_log01_cropping = mean(aphelenchoides_log01_cropping)
aphelenchoides_log02_cropping = mean(aphelenchoides_log02_cropping)

odd_aphelenchoides_cropping = c(aphelenchoides_log01_cropping,aphelenchoides_log02_cropping)


odd_aphelenchoides_cropping = as.data.frame(odd_aphelenchoides_cropping)
odd_aphelenchoides_cropping$nematode = c("Aphelenchoides","Aphelenchoides")
colnames(odd_aphelenchoides_cropping) = c("odd","nematode")
#odd_aphelenchoides_cropping$odd<-exp(odd_aphelenchoides_cropping$odd)

#----------------------------------------------------------

#Criconematidae

criconematidae_log$predicted = predict(crico_log, type = "response")
#criconematidae_log$predicted = exp(criconematidae_log$predicted)
criconematidae_log$predicted = (criconematidae_log$predicted/(1 - criconematidae_log$predicted))

criconematidae_log01_cropping<-criconematidae_log$predicted[criconematidae_log$cropping_corrigido==1]
criconematidae_log02_cropping<-criconematidae_log$predicted[criconematidae_log$cropping_corrigido==2]

criconematidae_log01_cropping = mean(criconematidae_log01_cropping)
criconematidae_log02_cropping = mean(criconematidae_log02_cropping)

odd_criconematidae_cropping = c(criconematidae_log01_cropping,criconematidae_log02_cropping)


odd_criconematidae_cropping = as.data.frame(odd_criconematidae_cropping)
odd_criconematidae_cropping$nematode = c("Criconematidae","Criconematidae")
colnames(odd_criconematidae_cropping) = c("odd","nematode")
#odd_criconematidae_cropping$odd<-exp(odd_criconematidae_cropping$odd)

#----------------------------------------------------------

#Pratylenchus

pratylenchus_log$predicted = predict(praty_log, type = "response")
pratylenchus_log$predicted = (pratylenchus_log$predicted/(1 - pratylenchus_log$predicted))


pratylenchus_log01_cropping<-pratylenchus_log$predicted[pratylenchus_log$cropping_corrigido==1]
pratylenchus_log02_cropping<-pratylenchus_log$predicted[pratylenchus_log$cropping_corrigido==2]

pratylenchus_log01_cropping = mean(pratylenchus_log01_cropping)
pratylenchus_log02_cropping = mean(pratylenchus_log02_cropping)

odd_pratylenchus_cropping = c(pratylenchus_log01_cropping,pratylenchus_log02_cropping)


odd_pratylenchus_cropping = as.data.frame(odd_pratylenchus_cropping)
odd_pratylenchus_cropping$nematode = c("Pratylenchus","Pratylenchus")
colnames(odd_pratylenchus_cropping) = c("odd","nematode")
#odd_pratylenchus_cropping$odd<-exp(odd_pratylenchus_cropping$odd)

#----------------------------------------------------------

#Rotylenchulus

rotylenchulus_log$predicted = predict(rotyl_log, type = "response")
#rotylenchulus_log$predicted = exp(rotylenchulus_log$predicted)
rotylenchulus_log$predicted = (rotylenchulus_log$predicted/(1 - rotylenchulus_log$predicted))

rotylenchulus_log01_cropping<-rotylenchulus_log$predicted[rotylenchulus_log$cropping_corrigido==1]
rotylenchulus_log02_cropping<-rotylenchulus_log$predicted[rotylenchulus_log$cropping_corrigido==2]

rotylenchulus_log01_cropping = mean(rotylenchulus_log01_cropping)
rotylenchulus_log02_cropping = mean(rotylenchulus_log02_cropping)

odd_rotylenchulus_cropping = c(rotylenchulus_log01_cropping,rotylenchulus_log02_cropping)


odd_rotylenchulus_cropping = as.data.frame(odd_rotylenchulus_cropping)
odd_rotylenchulus_cropping$nematode = c("Rotylenchulus","Rotylenchulus")
colnames(odd_rotylenchulus_cropping) = c("odd","nematode")
#odd_rotylenchulus_cropping$odd = exp(odd_rotylenchulus_cropping$odd)


all_odd_cropping = rbind(odd_aphelenchoides_cropping,odd_aphelenchus_cropping,
                         odd_criconematidae_cropping,
                         odd_helicotylenchus_cropping,odd_pratylenchus_cropping,
                         odd_rotylenchulus_cropping)
  #all_odd_cropping$odd = exp(all_odd_cropping$odd)

#----------------------------------------------------------
  
odd_ratio_cropping = all_odd_cropping %>% 
    group_by(nematode) %>% 
  summarise(
    n = n(),
    ratio = odd[1]/odd[2],
    #ratio_exp = exp(odd[1]/odd[2]),
    mean = mean(odd),
    sd = sd(odd),
   se=sd/sqrt(n))%>% 
  group_by(nematode) %>% 
   mutate(
    lower_log = log(ratio) - 1.96 * se, # Calcula o limite inferior do intervalo de confiança na escala logarítmica
    upper_log = log(ratio) + 1.96 * se, # Calcula o limite superior do intervalo de confiança na escala logarítmica
    lower = plogis(lower_log), # Transforma o limite inferior para a escala linear (odds ratio)
    upper = plogis(upper_log)  # Transforma o limite superior para a escala linear (odds ratio)
  )
  
  
  #mutate(
   #ic =se * qt((1-0.05)/2 + .5, n-1),
  # lower = (ratio_exp - 1.96 * se),
  # upper = (ratio_exp + 1.96 * se)
  # )

odd_ratio_cropping$practice = c("cropping","cropping","cropping","cropping",
                                "cropping","cropping")

#----------------------------------------------------------

cropping = odd_ratio_cropping %>% 
  #filter(!nematode %in% c("Helicotylenchus","Aphelenchoides")) %>% 
  ggplot(aes(nematode,ratio))+
  geom_hline(yintercept=1, linetype="dashed", color = "black", size = 1)+
  geom_pointrange(aes(ymin = ratio - lower, ymax = ratio + upper), color = "black", size = 1)+
    theme_minimal()+
  labs(x = "Morphological group",
       y = "Odd ratio")+
  theme(text = element_text(size = 22))+
  coord_flip()


#ggsave("odd_ratio_cropping.png", dpi = 600, bg = "white")

```

### By tillage

```{r}

#Helicotylenchus

helicotylenchus_log$predicted = predict(heli_log, type = "response")
helicotylenchus_log$predicted = (helicotylenchus_log$predicted/(1 - helicotylenchus_log$predicted))
#helicotylenchus_log$predicted = exp(helicotylenchus_log$predicted)
#helicotylenchus_log$predicted1 = predict(heli_log, type = "link")
#helicotylenchus_log$predicted1 = exp(helicotylenchus_log$predicted1)

helicotylenchus_log01_tillage<-helicotylenchus_log$predicted[helicotylenchus_log$tillage_corrigido==1]
helicotylenchus_log02_tillage<-helicotylenchus_log$predicted[helicotylenchus_log$tillage_corrigido==2]

helicotylenchus_log01_tillage = mean(helicotylenchus_log01_tillage)
helicotylenchus_log02_tillage = mean(helicotylenchus_log02_tillage)


odd_helicotylenchus_tillage = c(helicotylenchus_log01_tillage,helicotylenchus_log02_tillage)


#odd_helicotylenchus_tillage<-exp(odd_helicotylenchus_tillage)

odd_helicotylenchus_tillage = as.data.frame(odd_helicotylenchus_tillage)
odd_helicotylenchus_tillage$nematode = c("Helicotylenchus","Helicotylenchus")
colnames(odd_helicotylenchus_tillage) = c("odd","nematode")
#odd_helicotylenchus_tillage$odd = exp(odd_helicotylenchus_tillage$odd)

#----------------------------------------------------------

#Aphelenchus

aphelenchus_log$predicted = predict(aphel_log, type = "response")
#aphelenchus_log$predicted = exp(aphelenchus_log$predicted)

aphelenchus_log$predicted = (aphelenchus_log$predicted/(1 - aphelenchus_log$predicted))

aphelenchus_log01_tillage<-aphelenchus_log$predicted[aphelenchus_log$tillage_corrigido==1]
aphelenchus_log02_tillage<-aphelenchus_log$predicted[aphelenchus_log$tillage_corrigido==2]

aphelenchus_log01_tillage = mean(aphelenchus_log01_tillage)
aphelenchus_log02_tillage = mean(aphelenchus_log02_tillage)

odd_aphelenchus_tillage = c(aphelenchus_log01_tillage,aphelenchus_log02_tillage)


odd_aphelenchus_tillage = as.data.frame(odd_aphelenchus_tillage)
odd_aphelenchus_tillage$nematode = c("Aphelenchus","Aphelenchus")
colnames(odd_aphelenchus_tillage) = c("odd","nematode")
#odd_aphelenchus_tillage$odd = exp(odd_aphelenchus_tillage$odd)

#----------------------------------------------------------

#Aphelenchoides

aphelenchoides_log$predicted = predict(aphels_log, type = "response")
#aphelenchoides_log$predicted = exp(aphelenchoides_log$predicted)

aphelenchoides_log$predicted = (aphelenchoides_log$predicted/(1 - aphelenchoides_log$predicted))

aphelenchoides_log01_tillage<-aphelenchoides_log$predicted[aphelenchoides_log$tillage_corrigido==1]
aphelenchoides_log02_tillage<-aphelenchoides_log$predicted[aphelenchoides_log$tillage_corrigido==2]

aphelenchoides_log01_tillage = mean(aphelenchoides_log01_tillage)
aphelenchoides_log02_tillage = mean(aphelenchoides_log02_tillage)

odd_aphelenchoides_tillage = c(aphelenchoides_log01_tillage,aphelenchoides_log02_tillage)


odd_aphelenchoides_tillage = as.data.frame(odd_aphelenchoides_tillage)
odd_aphelenchoides_tillage$nematode = c("Aphelenchoides","Aphelenchoides")
colnames(odd_aphelenchoides_tillage) = c("odd","nematode")
#odd_aphelenchoides_tillage$odd = exp(odd_aphelenchoides_tillage$odd)

#----------------------------------------------------------

#Criconematidae

criconematidae_log$predicted = predict(crico_log, type = "response")
#criconematidae_log$predicted = exp(criconematidae_log$predicted)

criconematidae_log$predicted = (criconematidae_log$predicted/(1 - criconematidae_log$predicted))

criconematidae_log01_tillage<-criconematidae_log$predicted[criconematidae_log$tillage_corrigido==1]
criconematidae_log02_tillage<-criconematidae_log$predicted[criconematidae_log$tillage_corrigido==2]

criconematidae_log01_tillage = mean(criconematidae_log01_tillage)
criconematidae_log02_tillage = mean(criconematidae_log02_tillage)

odd_criconematidae_tillage = c(criconematidae_log01_tillage,criconematidae_log02_tillage)


odd_criconematidae_tillage = as.data.frame(odd_criconematidae_tillage)
odd_criconematidae_tillage$nematode = c("Criconematidae","Criconematidae")
colnames(odd_criconematidae_tillage) = c("odd","nematode")
#odd_criconematidae_tillage$odd = exp(odd_criconematidae_tillage$odd)

#----------------------------------------------------------

#Pratylenchus

pratylenchus_log$predicted = predict(praty_log, type = "response")
#pratylenchus_log$predicted = exp(pratylenchus_log$predicted)

pratylenchus_log$predicted = (pratylenchus_log$predicted/(1 - pratylenchus_log$predicted))

pratylenchus_log01_tillage<-pratylenchus_log$predicted[pratylenchus_log$tillage_corrigido==1]
pratylenchus_log02_tillage<-pratylenchus_log$predicted[pratylenchus_log$tillage_corrigido==2]

pratylenchus_log01_tillage = mean(pratylenchus_log01_tillage)
pratylenchus_log02_tillage = mean(pratylenchus_log02_tillage)

odd_pratylenchus_tillage = c(pratylenchus_log01_tillage,pratylenchus_log02_tillage)


odd_pratylenchus_tillage = as.data.frame(odd_pratylenchus_tillage)
odd_pratylenchus_tillage$nematode = c("Pratylenchus","Pratylenchus")
colnames(odd_pratylenchus_tillage) = c("odd","nematode")
#odd_pratylenchus_tillage$odd = exp(odd_pratylenchus_tillage$odd)

#----------------------------------------------------------

#Rotylenchulus

rotylenchulus_log$predicted = predict(rotyl_log, type = "response")
#rotylenchulus_log$predicted = exp(rotylenchulus_log$predicted)

rotylenchulus_log$predicted = (rotylenchulus_log$predicted/(1 - rotylenchulus_log$predicted))

rotylenchulus_log01_tillage<-rotylenchulus_log$predicted[rotylenchulus_log$tillage_corrigido==1]
rotylenchulus_log02_tillage<-rotylenchulus_log$predicted[rotylenchulus_log$tillage_corrigido==2]

rotylenchulus_log01_tillage = mean(rotylenchulus_log01_tillage)
rotylenchulus_log02_tillage = mean(rotylenchulus_log02_tillage)

odd_rotylenchulus_tillage = c(rotylenchulus_log01_tillage,rotylenchulus_log02_tillage)


odd_rotylenchulus_tillage = as.data.frame(odd_rotylenchulus_tillage)
odd_rotylenchulus_tillage$nematode = c("Rotylenchulus","Rotylenchulus")
colnames(odd_rotylenchulus_tillage) = c("odd","nematode")
#odd_rotylenchulus_tillage$odd = exp(odd_rotylenchulus_tillage$odd)



all_odd_tillage = rbind(odd_aphelenchoides_tillage,odd_aphelenchus_tillage,
                        odd_criconematidae_tillage,odd_helicotylenchus_tillage,
                        odd_pratylenchus_tillage,odd_rotylenchulus_tillage)
  #all_odd_tillage$odd = exp(all_odd_tillage$odd)

#----------------------------------------------------------

odd_ratio_tillage = all_odd_tillage %>% 
  group_by(nematode) %>% 
  summarise(
    n = n(),
    ratio = odd[1]/odd[2],
    #ratio_exp = exp(odd[1]/odd[2]),
    mean = mean(odd),
    sd = sd(odd),
   se=sd/sqrt(n))%>% 
  #group_by(nematode) %>% 
 mutate(
    lower_log = log(ratio) - 1.96 * se, # Calcula o limite inferior do intervalo de confiança na escala logarítmica
    upper_log = log(ratio) + 1.96 * se, # Calcula o limite superior do intervalo de confiança na escala logarítmica
    lower = plogis(lower_log), # Transforma o limite inferior para a escala linear (odds ratio)
    upper = plogis(upper_log)  # Transforma o limite superior para a escala linear (odds ratio)
  )



odd_ratio_tillage$practice = c("tillage","tillage","tillage","tillage",
                                "tillage","tillage")

#----------------------------------------------------------

tillage = odd_ratio_tillage %>% 
  #filter(!nematode %in% c("Aphelenchoides","Helicotylenchus")) %>% 
  ggplot(aes(nematode,ratio))+
  geom_hline(yintercept=1, linetype="dashed", color = "black", size = 1)+
  geom_pointrange(aes(ymin = ratio - lower, ymax = ratio + upper), color = "black", size = 1)+
   theme_minimal()+
  labs(x = "",
       y = "Odd ratio")+
   theme(text = element_text(size = 22))+
  coord_flip()

ggsave("odd_ratio_tillage.png", dpi = 600, bg = "white")
```

```{r}
plot_grid(cropping,tillage, labels = c("A", "B"), ncol = 2,
          label_size = 25)

ggsave("odd_ratio_cropping_tillage2.png", dpi = 600, bg = "white", width = 18, height = 10)
```

### By soil (2x6)

```{r}


#Helicotylenchus
helicotylenchus_log$predicted = predict(heli_log, type = "response")
helicotylenchus_log$predicted = (helicotylenchus_log$predicted/(1 - helicotylenchus_log$predicted))

helicotylenchus_log01<-helicotylenchus_log$predicted[helicotylenchus_log$soil_corrigido==2]
helicotylenchus_log02<-helicotylenchus_log$predicted[helicotylenchus_log$soil_corrigido==6]

helicotylenchus_log01 = mean(helicotylenchus_log01)
helicotylenchus_log02 = mean(helicotylenchus_log02)


odd_helicotylenchus = c(helicotylenchus_log01,helicotylenchus_log02)

#logodd_diff<-s1/s2
#odd_helicotylenchus<-exp(odd_helicotylenchus)

odd_helicotylenchus = as.data.frame(odd_helicotylenchus)
odd_helicotylenchus$nematode = c("Helicotylenchus","Helicotylenchus")
colnames(odd_helicotylenchus) = c("odd","nematode")

#----------------------------------------------------------

#Aphelenchus

aphelenchus_log$predicted = predict(aphel_log, type = "response")
aphelenchus_log$predicted = (aphelenchus_log$predicted/(1 - aphelenchus_log$predicted))

aphelenchus_log01<-aphelenchus_log$predicted[aphelenchus_log$soil_corrigido==2]
aphelenchus_log02<-aphelenchus_log$predicted[aphelenchus_log$soil_corrigido==6]

aphelenchus_log01 = mean(aphelenchus_log01)
aphelenchus_log02 = mean(aphelenchus_log02)

odd_aphelenchus = c(aphelenchus_log01,aphelenchus_log02)
#odd_aphelenchus<-exp(odd_aphelenchus)

odd_aphelenchus = as.data.frame(odd_aphelenchus)
odd_aphelenchus$nematode = c("Aphelenchus","Aphelenchus")
colnames(odd_aphelenchus) = c("odd","nematode")

#----------------------------------------------------------

#Aphelenchoides

aphelenchoides_log$predicted = predict(aphels_log, type = "response")
aphelenchoides_log$predicted = (aphelenchoides_log$predicted/(1 - aphelenchoides_log$predicted))

aphelenchoides_log01<-aphelenchoides_log$predicted[aphelenchoides_log$soil_corrigido==2]
aphelenchoides_log02<-aphelenchoides_log$predicted[aphelenchoides_log$soil_corrigido==6]

aphelenchoides_log01 = mean(aphelenchoides_log01)
aphelenchoides_log02 = mean(aphelenchoides_log02)

odd_aphelenchoides = c(aphelenchoides_log01,aphelenchoides_log02)
#odd_aphelenchoides<-exp(odd_aphelenchoides)

odd_aphelenchoides = as.data.frame(odd_aphelenchoides)
odd_aphelenchoides$nematode = c("Aphelenchoides","Aphelenchoides")
colnames(odd_aphelenchoides) = c("odd","nematode")

#----------------------------------------------------------

#Criconematidae

criconematidae_log$predicted = predict(crico_log, type = "response")
criconematidae_log$predicted = (criconematidae_log$predicted/(1 - criconematidae_log$predicted))

criconematidae_log01<-criconematidae_log$predicted[criconematidae_log$soil_corrigido==2]
criconematidae_log02<-criconematidae_log$predicted[criconematidae_log$soil_corrigido==6]

criconematidae_log01 = mean(criconematidae_log01)
criconematidae_log02 = mean(criconematidae_log02)

odd_criconematidae = c(criconematidae_log01,criconematidae_log02)
#odd_criconematidae<-exp(odd_criconematidae)

odd_criconematidae = as.data.frame(odd_criconematidae)
odd_criconematidae$nematode = c("Criconematidae","Criconematidae")
colnames(odd_criconematidae) = c("odd","nematode")

#----------------------------------------------------------

#Pratylenchus

pratylenchus_log$predicted = predict(praty_log, type = "response")
pratylenchus_log$predicted = (pratylenchus_log$predicted/(1 - pratylenchus_log$predicted))

pratylenchus_log01<-pratylenchus_log$predicted[pratylenchus_log$soil_corrigido==2]
pratylenchus_log02<-pratylenchus_log$predicted[pratylenchus_log$soil_corrigido==6]

pratylenchus_log01 = mean(pratylenchus_log01)
pratylenchus_log02 = mean(pratylenchus_log02)

odd_pratylenchus = c(pratylenchus_log01,pratylenchus_log02)
#odd_pratylenchus<-exp(odd_pratylenchus)

odd_pratylenchus = as.data.frame(odd_pratylenchus)
odd_pratylenchus$nematode = c("Pratylenchus","Pratylenchus")
colnames(odd_pratylenchus) = c("odd","nematode")

#----------------------------------------------------------

#Rotylenchulus

rotylenchulus_log$predicted = predict(rotyl_log, type = "response")
rotylenchulus_log$predicted = (rotylenchulus_log$predicted/(1 - rotylenchulus_log$predicted))

rotylenchulus_log01<-rotylenchulus_log$predicted[rotylenchulus_log$soil_corrigido==2]
rotylenchulus_log02<-rotylenchulus_log$predicted[rotylenchulus_log$soil_corrigido==6]

rotylenchulus_log01 = mean(rotylenchulus_log01)
rotylenchulus_log02 = mean(rotylenchulus_log02)

odd_rotylenchulus = c(rotylenchulus_log01,rotylenchulus_log02)
#odd_rotylenchulus<-exp(odd_rotylenchulus)

odd_rotylenchulus = as.data.frame(odd_rotylenchulus)
odd_rotylenchulus$nematode = c("Rotylenchulus","Rotylenchulus")
colnames(odd_rotylenchulus) = c("odd","nematode")

all_odd_02 = rbind(odd_aphelenchoides,odd_aphelenchus,odd_criconematidae,odd_helicotylenchus,
                odd_pratylenchus,odd_rotylenchulus)

#----------------------------------------------------------
  
odd_ratio_genus_02 = all_odd_02 %>% 
  group_by(nematode) %>% 
  summarise(
    n = n(),
    ratio = odd[1]/odd[2],
    mean = mean(odd),
    sd = sd(odd),
   se=sd/sqrt(n))%>% 
  #group_by(nematode) %>% 
  mutate(
    lower_log = log(ratio) - 1.96 * se, # Calcula o limite inferior do intervalo de confiança na escala logarítmica
    upper_log = log(ratio) + 1.96 * se, # Calcula o limite superior do intervalo de confiança na escala logarítmica
    lower = plogis(lower_log), # Transforma o limite inferior para a escala linear (odds ratio)
    upper = plogis(upper_log)  # Transforma o limite superior para a escala linear (odds ratio)
  )


odd_ratio_genus_02$soil_region = c("2x6","2x6","2x6","2x6","2x6","2x6")
odd_ratio_genus_02$soil_region = c("Argissolo|Latossolo","Argissolo|Latossolo",
                                   "Argissolo|Latossolo","Argissolo|Latossolo",
                                   "Argissolo|Latossolo","Argissolo|Latossolo")

#----------------------------------------------------------

ratio_02 = odd_ratio_genus_02 %>%  
 # filter(!nematode == "Rotylenchulus") %>% 
  ggplot(aes(nematode,ratio))+
  geom_hline(yintercept=1, linetype="dashed", color = "black", size = 1)+
  geom_pointrange(aes(ymin = ratio - lower, ymax = ratio + upper), color = "black", size = 0.7)+
  theme_minimal()+
  labs(x = "Morphological group",
       y = "Odd ratio")+
   theme(text = element_text(size = 22))+
  coord_flip()

  
```

### By soil (5x6)

```{r}

#Helicotylenchus
helicotylenchus_log$predicted = predict(heli_log, type = "response")
helicotylenchus_log$predicted = (helicotylenchus_log$predicted/(1 - helicotylenchus_log$predicted))

helicotylenchus_log01<-helicotylenchus_log$predicted[helicotylenchus_log$soil_corrigido==5]
helicotylenchus_log02<-helicotylenchus_log$predicted[helicotylenchus_log$soil_corrigido==6]

helicotylenchus_log01 = mean(helicotylenchus_log01)
helicotylenchus_log02 = mean(helicotylenchus_log02)


odd_helicotylenchus = c(helicotylenchus_log01,helicotylenchus_log02)

#odd_helicotylenchus<-exp(odd_helicotylenchus)

odd_helicotylenchus = as.data.frame(odd_helicotylenchus)
odd_helicotylenchus$nematode = c("Helicotylenchus","Helicotylenchus")
colnames(odd_helicotylenchus) = c("odd","nematode")

#----------------------------------------------------------

#Aphelenchus

aphelenchus_log$predicted = predict(aphel_log, type = "response")
aphelenchus_log$predicted = (aphelenchus_log$predicted/(1 - aphelenchus_log$predicted))

aphelenchus_log01<-aphelenchus_log$predicted[aphelenchus_log$soil_corrigido==5]
aphelenchus_log02<-aphelenchus_log$predicted[aphelenchus_log$soil_corrigido==6]

aphelenchus_log01 = mean(aphelenchus_log01)
aphelenchus_log02 = mean(aphelenchus_log02)

odd_aphelenchus = c(aphelenchus_log01,aphelenchus_log02)
#odd_aphelenchus<-exp(odd_aphelenchus)

odd_aphelenchus = as.data.frame(odd_aphelenchus)
odd_aphelenchus$nematode = c("Aphelenchus","Aphelenchus")
colnames(odd_aphelenchus) = c("odd","nematode")

#----------------------------------------------------------

#Aphelenchoides

aphelenchoides_log$predicted = predict(aphels_log, type = "response")
aphelenchoides_log$predicted = (aphelenchoides_log$predicted/(1 - aphelenchoides_log$predicted))

aphelenchoides_log01<-aphelenchoides_log$predicted[aphelenchoides_log$soil_corrigido==5]
aphelenchoides_log02<-aphelenchoides_log$predicted[aphelenchoides_log$soil_corrigido==6]

aphelenchoides_log01 = mean(aphelenchoides_log01)
aphelenchoides_log02 = mean(aphelenchoides_log02)

odd_aphelenchoides = c(aphelenchoides_log01,aphelenchoides_log02)
#odd_aphelenchoides<-exp(odd_aphelenchoides)

odd_aphelenchoides = as.data.frame(odd_aphelenchoides)
odd_aphelenchoides$nematode = c("Aphelenchoides","Aphelenchoides")
colnames(odd_aphelenchoides) = c("odd","nematode")

#----------------------------------------------------------

#Criconematidae

criconematidae_log$predicted = predict(crico_log, type = "response")
criconematidae_log$predicted = (criconematidae_log$predicted/(1 - criconematidae_log$predicted))

criconematidae_log01<-criconematidae_log$predicted[criconematidae_log$soil_corrigido==5]
criconematidae_log02<-criconematidae_log$predicted[criconematidae_log$soil_corrigido==6]

criconematidae_log01 = mean(criconematidae_log01)
criconematidae_log02 = mean(criconematidae_log02)

odd_criconematidae = c(criconematidae_log01,criconematidae_log02)
#odd_criconematidae<-exp(odd_criconematidae)

odd_criconematidae = as.data.frame(odd_criconematidae)
odd_criconematidae$nematode = c("Criconematidae","Criconematidae")
colnames(odd_criconematidae) = c("odd","nematode")

#----------------------------------------------------------

#Pratylenchus

pratylenchus_log$predicted = predict(praty_log, type = "response")
pratylenchus_log$predicted = (pratylenchus_log$predicted/(1 - pratylenchus_log$predicted))

pratylenchus_log01<-pratylenchus_log$predicted[pratylenchus_log$soil_corrigido==5]
pratylenchus_log02<-pratylenchus_log$predicted[pratylenchus_log$soil_corrigido==6]

pratylenchus_log01 = mean(pratylenchus_log01)
pratylenchus_log02 = mean(pratylenchus_log02)

odd_pratylenchus = c(pratylenchus_log01,pratylenchus_log02)
#odd_pratylenchus<-exp(odd_pratylenchus)

odd_pratylenchus = as.data.frame(odd_pratylenchus)
odd_pratylenchus$nematode = c("Pratylenchus","Pratylenchus")
colnames(odd_pratylenchus) = c("odd","nematode")

#----------------------------------------------------------

#Rotylenchulus

rotylenchulus_log$predicted = predict(rotyl_log, type = "response")
rotylenchulus_log$predicted = (rotylenchulus_log$predicted/(1 - rotylenchulus_log$predicted))

rotylenchulus_log01<-rotylenchulus_log$predicted[rotylenchulus_log$soil_corrigido==5]
rotylenchulus_log02<-rotylenchulus_log$predicted[rotylenchulus_log$soil_corrigido==6]

rotylenchulus_log01 = mean(rotylenchulus_log01)
rotylenchulus_log02 = mean(rotylenchulus_log02)

odd_rotylenchulus = c(rotylenchulus_log01,rotylenchulus_log02)
#odd_rotylenchulus<-exp(odd_rotylenchulus)

odd_rotylenchulus = as.data.frame(odd_rotylenchulus)
odd_rotylenchulus$nematode = c("Rotylenchulus","Rotylenchulus")
colnames(odd_rotylenchulus) = c("odd","nematode")

all_odd_05 = rbind(odd_aphelenchoides,odd_aphelenchus,odd_criconematidae,odd_helicotylenchus,
                odd_pratylenchus,odd_rotylenchulus)

#----------------------------------------------------------
  
odd_ratio_genus_05 = all_odd_05 %>% 
  group_by(nematode) %>% 
 summarise(
    n = n(),
    ratio = odd[1]/odd[2],
    ratio_exp = exp(odd[1]/odd[2]),
    mean = mean(odd),
    sd = sd(odd),
   se=sd/sqrt(n))%>% 
  #group_by(nematode) %>% 
    mutate(
    lower_log = log(ratio) - 1.96 * se, # Calcula o limite inferior do intervalo de confiança na escala logarítmica
    upper_log = log(ratio) + 1.96 * se, # Calcula o limite superior do intervalo de confiança na escala logarítmica
    lower = plogis(lower_log), # Transforma o limite inferior para a escala linear (odds ratio)
    upper = plogis(upper_log)  # Transforma o limite superior para a escala linear (odds ratio)
  )

odd_ratio_genus_05$soil_region = c("5x6","5x6","5x6","5x6","5x6","5x6")
odd_ratio_genus_05$soil_region = c("Cambissolo|Latossolo","Cambissolo|Latossolo","Cambissolo|Latossolo",
                                   "Cambissolo|Latossolo","Cambissolo|Latossolo",
                                   "Cambissolo|Latossolo") #comparação - thais

#----------------------------------------------------------

ratio_05 = odd_ratio_genus_05 %>% 
 # filter(!nematode == "Rotylenchulus") %>% 
  ggplot(aes(nematode,ratio))+
  geom_hline(yintercept=1, linetype="dashed", color = "black", size = 1)+
  geom_pointrange(aes(ymin = ratio - lower, ymax = ratio + upper), color = "black", size = 0.7)+
  theme_minimal()+
  labs(x = "",
       y = "Odd ratio")+
  theme(text = element_text(size = 22))+
        #axis.text.y = element_blank())+
    coord_flip()

```

### Plotting

```{r}
 plot_grid(ratio_02,ratio_05, labels = c("A", "B"), ncol = 2,
          label_size = 25)

ggsave("odd_ratio_soil_mixed.png",dpi = 600, bg = "white", width = 18, height = 10)
```

## Altitude

```{r}

helicotylenchus_log$prob = predict(heli_log, type = "response")
criconematidae_log$prob = predict(crico_log, type = "response")
aphelenchus_log$prob = predict(aphel_log, type = "response")
aphelenchoides_log$prob = predict(aphels_log, type = "response")
pratylenchus_log$prob = predict(praty_log, type = "response")
rotylenchulus_log$prob = predict(rotyl_log, type = "response")
#exp(summary(heli_log))


altitude = rbind(helicotylenchus_log,criconematidae_log,aphelenchus_log,aphelenchoides_log,
                 pratylenchus_log,rotylenchulus_log)
 altitude %>% 
   filter(altitute < 400)


altitude1 <- altitude %>%
  mutate(altitute = case_when(altitute >= 0 ~ altitute,
                                altitute == NA_character_ ~ 0 )) %>% 
  mutate(altitude_range = case_when(
    altitute <= 400 ~ "280 to 400",
    altitute <= 600  ~ "401 to 600",
    altitute <= 800 ~ "601 to 800",
    altitute <= 1000 ~ "801 to 1000",
    altitute > 1000 ~"Above 1001"
  ))

altitude_genus = altitude1 %>% 
    group_by(genus,altitude_range) %>% 
  summarise(
    
    prob = prob)
   


point = altitude1 %>% 
    group_by(genus,altitude_range) %>% 
  summarise(
    
    probi = mean(prob),
    sd = sd(prob)
   
  ) %>%
  group_by(genus,altitude_range) %>% 
  summarise(
   prob = mean(probi),
   sd = sd
   )
 
#write_xlsx(point,"THAIS/point.xlsx")


point %>% 
    ggplot(aes(altitude_range,prob))+
  geom_jitter(data = altitude_genus, alpha = .2, size = 5, width = .03, color = "gray")+
  #geom_violin(data = altitude_genus, alpha = .7, width = .03, fill = "grey", color = "grey")+
  geom_pointrange(aes(ymin = prob- sd, ymax = prob + sd), color = "black", size = 1)+
  theme_minimal()+
  facet_wrap(~genus)+
  labs(x= "Range altitude (m)",
       y = "Probability")+
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 45, hjust = 1))
  
#ggsave("prob_altitude.png", bg = "white", width = 12, height = 8)
```

## Overall

```{r}
altitude_genus = altitude1 %>% 
    group_by(genus,altitude_range,sand,clay,ph,cropping,soil,tillage) %>% 
  summarise(
    
    prob = prob
   
  )

#write_xlsx(altitude_genus,"THAIS/altitude_genus.xlsx")

```

## Probability of incidence

### By tillage and sand

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+tillage, nrow = 3)+
  theme_minimal()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_sand_tillage_all_genus.png", dpi = 600, width = 14, height = 8, bg = "white")
```

### By tillage and  pH

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(ph,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+tillage, nrow = 3)+
  theme_minimal()+
  labs(x = "pH gradient",
       y = "Probabilitiy")+
  theme(text = element_text(size = 18))

ggsave("probability_ph_tillage_all_genus.png", dpi = 600, width = 14, height = 8,bg = "white")
```

### By tillage and clay

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(clay,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+tillage, nrow = 3)+
  theme_minimal()+
  labs(x = "Clay",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_clay_tillage_all_genus.png", dpi = 600, width = 14, height = 8, bg = "white")
```

### By sand and soil type

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F,color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+soil, nrow = 3)+
  theme_minimal()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_soil_all_genus.png", dpi = 600, width = 14,height = 8, bg = "white")
```

### By cropping and sand

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+cropping, nrow = 3)+
  theme_minimal()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_cropping_sand_all_genus.png", dpi = 600, width = 14, height = 8, bg = "white")
```

### By cropping and pH

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(ph,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black",fullrange = T)+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+cropping, nrow = 3)+
  theme_minimal()+
  labs(x = "pH gradient",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_cropping_ph_all_genus.png", dpi = 600, width = 14,height = 6, bg = "white")
```

### By sand

```{r}
altitude_genus %>% 
  #filter(!genus %in% c("Aphelenchus","Criconematidae","Rotylenchulus")) %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 2)+
  theme_minimal()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_sand_genus.png", dpi = 600, width = 14, height = 6, bg = "white")
```

### By pH

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(ph,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 2)+
  theme_minimal()+
  labs(x = "pH",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_ph_genus.png", dpi = 600, width = 14,height = 6, bg = "white")
```

### By clay

```{r}
altitude_genus %>% 
 #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(clay,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 2)+
  theme_minimal()+
  labs(x = "Clay",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_clay_genus.png", dpi = 600, width = 14,height = 6, bg = "white")
```


# Abundance - Ordinal regression

```{r}
#Modificado
population = gsheet2tbl("https://docs.google.com/spreadsheets/d/1vtTIBgpJ1qqdPp3MjIPUEYASMj_lUSZR/edit?usp=sharing&ouid=112586075609758894128&rtpof=true&sd=true")

 population$quantity = 6.66*population$quantity 
 population$quantity =as.integer(population$quantity)
 population$quantity =as.numeric(population$quantity)

#population[,10] = NULL
 
```


## By Helicotylenchus

```{r}
#library(MASS)
library(ordinal)

helicotylenchus_ordinal = population %>% 
  filter(genus == "Helicotylenchus") %>%
  filter(soil %in% c("2","5","6")) %>% 
   group_by(samples,soil,altitute,sand,clay,ph,ec,cropping,tillage) %>% 
  summarise(
    quantity = sum(quantity)
  )

helicotylenchus_ordinal$ph = as.numeric(helicotylenchus_ordinal$ph)
helicotylenchus_ordinal <- helicotylenchus_ordinal %>%
  mutate(quantity = case_when(quantity >= 0 ~ quantity,
                                quantity == NA_character_ ~ 0 )) %>% 
  mutate(risk_category = case_when(
    quantity == 0 ~ "0",
    quantity <= 75  ~ "1",
    quantity <= 150 ~ "2",
    quantity <= 300 ~ "2",
    quantity <= 500 ~"2",
    quantity > 500 ~"2"
  ))

helicotylenchus_ordinal$risk_category = as.factor(helicotylenchus_ordinal$risk_category)
#helicotylenchus_ordinal$tillage = as.character(helicotylenchus_ordinal$tillage)
#helicotylenchus_ordinal$soil = as.character(helicotylenchus_ordinal$soil)
#helicotylenchus_ordinal$cropping = as.character(helicotylenchus_ordinal$cropping)

helic_ordinal = clm(risk_category ~ tillage + cropping + soil + sand + clay  +
                       ph,
                     data = helicotylenchus_ordinal,link = "logit")


helic_predicted =cbind(helicotylenchus_ordinal, predict(helic_ordinal, helicotylenchus_ordinal, type = "prob")$fit)
colnames(helic_predicted) = c("samples","soil","altitute","sand","clay","ph","ec",
                              "cropping","tillage","quantity","risk_category",
                              "prob")
#Summary


coef(summary(helic_ordinal))
coef(helic_ordinal)

#Odd calculate
helic_predicted = helic_predicted %>% 
  group_by(samples,soil,altitute,sand,clay,ph,cropping,tillage,quantity,
         risk_category,prob) %>% 
  summarise(
    odd = (prob/(1-prob)
  ))

anova(helic_ordinal, type="III")
confint(helic_ordinal, type = c("profile"),level = .95)

#Wald X²

waldtest(helic_ordinal, term = 1)
waldtest(helic_ordinal, term = 2)
waldtest(helic_ordinal, term = 3)
waldtest(helic_ordinal, term = 4)
waldtest(helic_ordinal, term = 5)
waldtest(helic_ordinal, term = 6)
#waldtest(helic_ordinal, term = 7)


#LRT

lrtest(helic_ordinal)



#plogis(helic_ordinal)


#Coefficients

coef(summary(helic_ordinal))

```

### Cross-validation

```{r}
library(ordinalNet)

risk_helicotylenchus = helicotylenchus_ordinal$risk_category
class(risk_helicotylenchus)

helicotylenchus_ordinal_cv = helicotylenchus_ordinal

helicotylenchus_ordinal_cv["samples"] = NULL
helicotylenchus_ordinal_cv["ec"] = NULL
helicotylenchus_ordinal_cv["altitute"] = NULL
helicotylenchus_ordinal_cv["quantity"] = NULL
helicotylenchus_ordinal_cv["risk_category"] = NULL

x_ordinal_helicotylenchus = helicotylenchus_ordinal_cv
x_ordinal_helicotylenchus = as.matrix(x_ordinal_helicotylenchus)
class(x_ordinal_helicotylenchus)

helicotylenchus_cv = ordinalNetCV(x_ordinal_helicotylenchus, risk_helicotylenchus,
                                  tuneMethod=c("cvMisclass"),family = "cumulative", link = "logit",
                                  lambdaMinRatio = 1e-04,printProgress = FALSE)
summary(helicotylenchus_cv)
colMeans(summary(helicotylenchus_cv))

```

Helicotylenchus model diagnostic (NÃO ESTÁ AJUSTANDO IGUAL O DE INCIDÊNCIA)

```{r,eval = FALSE, echo = TRUE}
#Model log with best
helic_ordinal = clm(risk_category ~ tillage + cropping + soil + sand + clay  +
                       ph,
                     data = helicotylenchus_ordinal,link = "logit")


summary(helic_ordinal)

helic_predicted1 = helic_predicted |> 
  filter(!risk_category == 2)

pred_helicotylenchus_ordinal <- helic_predicted1$prob


  


#Better thresholds
dat_helicotylenchus <- data.frame(1,helic_predicted1$risk_category, pred_helicotylenchus_ordinal)
dat_helicotylenchus[dat_helicotylenchus$helic_predicted1.risk_category== "0",c("risk")] <- 0
dat_helicotylenchus[dat_helicotylenchus$helic_predicted1.risk_category== "1",c("risk")] <-1
dat_helicotylenchus["helic_predicted1.risk_category"] = NULL

optimal.thresholds(dat_helicotylenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_helicotylenchus_ordinal > 0.9)),  mode= "everything",  reference = as.factor(helicotylenchus_log$incidence))


#Plot
auc.roc.plot(dat_helicotylenchus)
calibration.plot(dat_helicotylenchus)

error.threshold.plot(dat_helicotylenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_helicotylenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_helicotylenchus,threshold=.9,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_helicotylenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_helicotylenchus <- roc.plot.calculate(dat_helicotylenchus) #data frame

roc_helicotylenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_helicotylenchus <- NULL
for(i in 1:nrow(helicotylenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- helicotylenchus_log[-i,]
    test <- helicotylenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = helicotylenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.9,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_helicotylenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_helicotylenchus)



```

### Odd ratio

```{r}

#Soil 2 x 6

helicotylenchus_log01_ordinal<-helic_predicted$odd[helic_predicted$soil==2]
helicotylenchus_log02_ordinal<-helic_predicted$odd[helic_predicted$soil==6]

helicotylenchus_log01_ordinal = mean(helicotylenchus_log01_ordinal)
helicotylenchus_log02_ordinal = mean(helicotylenchus_log02_ordinal)


odd_helicotylenchus_ordinal= c(helicotylenchus_log01_ordinal,helicotylenchus_log02_ordinal)
#odd_helicotylenchus_ordinal = exp(odd_helicotylenchus_ordinal)

odd_helicotylenchus_ordinal = as.data.frame(odd_helicotylenchus_ordinal)
odd_helicotylenchus_ordinal$nematode = c("Helicotylenchus","Helicotylenchus")
odd_helicotylenchus_ordinal$comparation = c("2x6","2x6")
colnames(odd_helicotylenchus_ordinal) = c("odd","nematode","comparation")

soil02_helicotylenchus = odd_helicotylenchus_ordinal


#Soil 5x6

helicotylenchus_log01_ordinal<-helic_predicted$odd[helic_predicted$soil==5]
helicotylenchus_log02_ordinal<-helic_predicted$odd[helic_predicted$soil==6]

helicotylenchus_log01_ordinal = mean(helicotylenchus_log01_ordinal)
helicotylenchus_log02_ordinal = mean(helicotylenchus_log02_ordinal)


odd_helicotylenchus_ordinal= c(helicotylenchus_log01_ordinal,helicotylenchus_log02_ordinal)
#odd_helicotylenchus_ordinal = exp(odd_helicotylenchus_ordinal)

odd_helicotylenchus_ordinal = as.data.frame(odd_helicotylenchus_ordinal)
odd_helicotylenchus_ordinal$nematode = c("Helicotylenchus","Helicotylenchus")
odd_helicotylenchus_ordinal$comparation = c("5x6","5x6")
colnames(odd_helicotylenchus_ordinal) = c("odd","nematode","comparation")

soil05_helicotylenchus = odd_helicotylenchus_ordinal

#Tillage 1x2

helicotylenchus_log01_ordinal<-helic_predicted$odd[helic_predicted$tillage==1]
helicotylenchus_log02_ordinal<-helic_predicted$odd[helic_predicted$tillage==2]

helicotylenchus_log01_ordinal = mean(helicotylenchus_log01_ordinal)
helicotylenchus_log02_ordinal = mean(helicotylenchus_log02_ordinal)


odd_helicotylenchus_ordinal= c(helicotylenchus_log01_ordinal,helicotylenchus_log02_ordinal)
#odd_helicotylenchus_ordinal = exp(odd_helicotylenchus_ordinal)

odd_helicotylenchus_ordinal = as.data.frame(odd_helicotylenchus_ordinal)
odd_helicotylenchus_ordinal$nematode = c("Helicotylenchus","Helicotylenchus")
odd_helicotylenchus_ordinal$comparation = c("tillage","tillage")
colnames(odd_helicotylenchus_ordinal) = c("odd","nematode","comparation")

tillage_helicotylenchus = odd_helicotylenchus_ordinal

#Cropping 1x2

helicotylenchus_log01_ordinal<-helic_predicted$odd[helic_predicted$cropping==1]
helicotylenchus_log02_ordinal<-helic_predicted$odd[helic_predicted$cropping==2]

helicotylenchus_log01_ordinal = mean(helicotylenchus_log01_ordinal)
helicotylenchus_log02_ordinal = mean(helicotylenchus_log02_ordinal)


odd_helicotylenchus_ordinal= c(helicotylenchus_log01_ordinal,helicotylenchus_log02_ordinal)
#odd_helicotylenchus_ordinal = exp(odd_helicotylenchus_ordinal)

odd_helicotylenchus_ordinal = as.data.frame(odd_helicotylenchus_ordinal)
odd_helicotylenchus_ordinal$nematode = c("Helicotylenchus","Helicotylenchus")
odd_helicotylenchus_ordinal$comparation = c("cropping","cropping")
colnames(odd_helicotylenchus_ordinal) = c("odd","nematode","comparation")

cropping_helicotylenchus = odd_helicotylenchus_ordinal

all_odd_helicotylenchus= rbind(soil02_helicotylenchus,soil05_helicotylenchus,
                               cropping_helicotylenchus,
                               tillage_helicotylenchus)

all_odd_ratio_helicotylenchus = all_odd_helicotylenchus %>% 
  group_by(comparation) %>% 
    summarise(
    n = n(),
    ratio = odd[1]/odd[2],
    ratio_exp =  exp(odd[1]/odd[2]),
    mean = mean(odd),
    sd = sd(odd),
   se=sd/sqrt(n)) %>% 
  group_by(comparation) %>% 
  mutate(
   lower = (ratio - 1.96 * se),
   upper = (ratio + 1.96 * se))

coef(summary(helic_ordinal))
```

## By Pratylenchus

```{r}
#Pratylenchus

pratylenchus_ordinal = population %>% 
  filter(genus == "Pratylenchus") %>% 
  filter(soil %in% c("2","5","6")) %>% 
   group_by(samples,soil,altitute,sand,clay,ph,ec,cropping,tillage) %>% 
  summarise(
    quantity = sum(quantity)
  )
pratylenchus_ordinal$ph = as.numeric(pratylenchus_ordinal$ph)

pratylenchus_ordinal <- pratylenchus_ordinal %>%
  mutate(quantity = case_when(quantity >= 0 ~ quantity,
                                quantity == NA_character_ ~ 0 )) %>% 
  mutate(risk_category = case_when(
    quantity == 0 ~ "0",
    quantity <= 10  ~ "1",
    quantity <= 25 ~ "2",
    quantity <= 50 ~ "2",
    quantity <= 100 ~"2",
    quantity > 100 ~ "2"
  ))

pratylenchus_ordinal$risk_category = as.factor(pratylenchus_ordinal$risk_category)
#pratylenchus_ordinal$tillage = as.character(pratylenchus_ordinal$tillage)
#pratylenchus_ordinal$soil = as.character(pratylenchus_ordinal$soil)
#pratylenchus_ordinal$cropping = as.character(pratylenchus_ordinal$cropping)

praty_ordinal = clm(risk_category ~ tillage + cropping + soil + sand + clay  +
                       ph,
                     data = pratylenchus_ordinal,link = "logit")



praty_predicted =cbind(pratylenchus_ordinal, predict(praty_ordinal, pratylenchus_ordinal, type = "prob")$fit)
colnames(praty_predicted) = c("samples","soil","altitute","sand","clay","ph","ec",
                              "cropping","tillage","quantity","risk_category",
                              "prob")


#Summary

exp(coef(summary(praty_ordinal)))
coef(summary(praty_ordinal))
coef(praty_ordinal)

anova(praty_ordinal, type="III")
confint(praty_ordinal, type = c("profile"),level = .95)


#Odd

praty_predicted = praty_predicted %>% 
  group_by(samples,soil,altitute,sand,clay,ph,cropping,tillage,quantity,
         risk_category,prob) %>% 
  summarise(
    odd = (prob/(1-prob)
  ))

#Wald X²

waldtest(praty_ordinal, term = 1)
waldtest(praty_ordinal, term = 2)
waldtest(praty_ordinal, term = 3)
waldtest(praty_ordinal, term = 4)
waldtest(praty_ordinal, term = 5)
waldtest(praty_ordinal, term = 6)
#waldtest(praty_ordinal, term = 7)
#waldtest(praty_ordinal, term = 8)

#LRT

lrtest(praty_ordinal)

```

### Cross-validation

```{r}

risk_pratylenchus = pratylenchus_ordinal$risk_category
class(risk_pratylenchus)

pratylenchus_ordinal_cv = pratylenchus_ordinal

pratylenchus_ordinal_cv["samples"] = NULL
pratylenchus_ordinal_cv["ec"] = NULL
pratylenchus_ordinal_cv["altitute"] = NULL
pratylenchus_ordinal_cv["quantity"] = NULL
pratylenchus_ordinal_cv["risk_category"] = NULL

x_ordinal_pratylenchus = pratylenchus_ordinal_cv
x_ordinal_pratylenchus = as.matrix(x_ordinal_pratylenchus)
class(x_ordinal_pratylenchus)

pratylenchus_cv = ordinalNetCV(x_ordinal_pratylenchus, risk_pratylenchus,family = "cumulative", link = "logit",
                                  tuneMethod=c("cvMisclass"),lambdaMinRatio = 1e-04,printProgress = FALSE)
summary(pratylenchus_cv)
colMeans(summary(pratylenchus_cv))

```

### Odd ratio

```{r}

#Soil 2 x 6

pratylenchus_log01_ordinal<-praty_predicted$odd[praty_predicted$soil==2]
pratylenchus_log02_ordinal<-praty_predicted$odd[praty_predicted$soil==6]

pratylenchus_log01_ordinal = mean(pratylenchus_log01_ordinal)
pratylenchus_log02_ordinal = mean(pratylenchus_log02_ordinal)


odd_pratylenchus_ordinal= c(pratylenchus_log01_ordinal,pratylenchus_log02_ordinal)

odd_pratylenchus_ordinal = as.data.frame(odd_pratylenchus_ordinal)
odd_pratylenchus_ordinal$nematode = c("Pratylenchus","Pratylenchus")
odd_pratylenchus_ordinal$comparation = c("2x6","2x6")
colnames(odd_pratylenchus_ordinal) = c("odd","nematode","comparation")

soil02_pratylenchus = odd_pratylenchus_ordinal


#Soil 5x6

pratylenchus_log01_ordinal<-praty_predicted$odd[praty_predicted$soil==5]
pratylenchus_log02_ordinal<-praty_predicted$odd[praty_predicted$soil==6]

pratylenchus_log01_ordinal = mean(pratylenchus_log01_ordinal)
pratylenchus_log02_ordinal = mean(pratylenchus_log02_ordinal)


odd_pratylenchus_ordinal= c(pratylenchus_log01_ordinal,pratylenchus_log02_ordinal)

odd_pratylenchus_ordinal = as.data.frame(odd_pratylenchus_ordinal)
odd_pratylenchus_ordinal$nematode = c("Pratylenchus","Pratylenchus")
odd_pratylenchus_ordinal$comparation = c("5x6","5x6")
colnames(odd_pratylenchus_ordinal) = c("odd","nematode","comparation")

soil05_pratylenchus = odd_pratylenchus_ordinal

#Tillage 1x2

pratylenchus_log01_ordinal<-praty_predicted$odd[praty_predicted$tillage==1]
pratylenchus_log02_ordinal<-praty_predicted$odd[praty_predicted$tillage==2]

pratylenchus_log01_ordinal = mean(pratylenchus_log01_ordinal)
pratylenchus_log02_ordinal = mean(pratylenchus_log02_ordinal)


odd_pratylenchus_ordinal= c(pratylenchus_log01_ordinal,pratylenchus_log02_ordinal)

odd_pratylenchus_ordinal = as.data.frame(odd_pratylenchus_ordinal)
odd_pratylenchus_ordinal$nematode = c("Pratylenchus","Pratylenchus")
odd_pratylenchus_ordinal$comparation = c("tillage","tillage")
colnames(odd_pratylenchus_ordinal) = c("odd","nematode","comparation")

tillage_pratylenchus = odd_pratylenchus_ordinal

#Cropping 1x2

pratylenchus_log01_ordinal<-praty_predicted$odd[praty_predicted$cropping==1]
pratylenchus_log02_ordinal<-praty_predicted$odd[praty_predicted$cropping==2]

pratylenchus_log01_ordinal = mean(pratylenchus_log01_ordinal)
pratylenchus_log02_ordinal = mean(pratylenchus_log02_ordinal)


odd_pratylenchus_ordinal= c(pratylenchus_log01_ordinal,pratylenchus_log02_ordinal)

odd_pratylenchus_ordinal = as.data.frame(odd_pratylenchus_ordinal)
odd_pratylenchus_ordinal$nematode = c("Pratylenchus","Pratylenchus")
odd_pratylenchus_ordinal$comparation = c("cropping","cropping")
colnames(odd_pratylenchus_ordinal) = c("odd","nematode","comparation")

cropping_pratylenchus = odd_pratylenchus_ordinal

all_odd_pratylenchus = rbind(soil02_pratylenchus,soil05_pratylenchus,
                             cropping_pratylenchus,
                       tillage_pratylenchus)

all_odd_ratio_pratylenchus = all_odd_pratylenchus %>% 
  group_by(comparation) %>% 
    summarise(
    n = n(),
    ratio = odd[1]/odd[2],
    mean = mean(odd),
    sd = sd(odd),
   se=sd/sqrt(n)) %>% 
  group_by(comparation) %>% 
  mutate(
   lower = (ratio - 1.96 * se),
   upper = (ratio + 1.96 * se))

coef(summary(praty_ordinal))
```

## By Criconematidae

```{r}
#Criconematidae

criconematidae_ordinal = population %>% 
  filter(genus == "Criconematidae") %>% 
  filter(soil %in% c("2","5","6")) %>% 
   group_by(samples,soil,altitute, sand,clay,ph,ec,cropping,tillage) %>% 
  summarise(
    quantity = sum(quantity)
  )
criconematidae_ordinal$ph = as.numeric(criconematidae_ordinal$ph)
criconematidae_ordinal <- criconematidae_ordinal %>%
  mutate(quantity = case_when(quantity >= 0 ~ quantity,
                                quantity == NA_character_ ~ 0 )) %>% 
  mutate(risk_category = case_when(
    quantity == 0 ~ "0",
    quantity <= 75  ~ "1",
    quantity <= 150 ~ "2",
    quantity <= 300 ~ "2",
    quantity < 600 ~"2",
    quantity >= 600 ~ "2"
  ))

criconematidae_ordinal$risk_category = as.factor(criconematidae_ordinal$risk_category)
#criconematidae_ordinal$tillage = as.character(criconematidae_ordinal$tillage)
#criconematidae_ordinal$soil = as.character(criconematidae_ordinal$soil)
#criconematidae_ordinal$cropping = as.character(criconematidae_ordinal$cropping)

crico_ordinal = clm(risk_category ~ tillage + cropping + soil + sand + clay  +
                       ph,
                     data = criconematidae_ordinal,link = "logit")



crico_predicted =cbind(criconematidae_ordinal, predict(crico_ordinal, criconematidae_ordinal, type = "prob")$fit)
colnames(crico_predicted) = c("samples","soil","altitute","sand","clay","ph","ec",
                              "cropping","tillage","quantity","risk_category",
                              "prob")

crico_predicted = crico_predicted %>% 
  group_by(samples,soil,altitute,sand,clay,ph,cropping,tillage,quantity,
         risk_category,prob) %>% 
  summarise(
    odd = (prob/(1-prob)
  ))

#Summary

exp(coef(summary(crico_ordinal)))
coef(summary(crico_ordinal))
coef(crico_ordinal)

anova(crico_ordinal, type="III")
confint(crico_ordinal, type = c("profile"),level = .95)

#Wald X²
coef(crico_ordinal)
waldtest(crico_ordinal, term = 1)
waldtest(crico_ordinal, term = 2)
waldtest(crico_ordinal, term = 3)
waldtest(crico_ordinal, term = 4)
waldtest(crico_ordinal, term = 5)
waldtest(crico_ordinal, term = 6)
#waldtest(crico_ordinal, term = 7)
#waldtest(crico_ordinal, term = 8)

#LRT

lrtest(crico_ordinal)

library(ggeffects)

#coefficents
coef(summary(crico_ordinal))
ggpredictions_ols1 = ggpredict(crico_ordinal, terms = "clay")
ggpredictions_ols1

#ggpredictions_ols1$ph = factor(ggpredictions_ols1$ph) # CONFERIR


```

### Cross-validation

```{r}

risk_criconematidae = criconematidae_ordinal$risk_category
class(risk_criconematidae)

criconematidae_ordinal["samples"] = NULL
criconematidae_ordinal["ec"] = NULL
criconematidae_ordinal["altitute"] = NULL
criconematidae_ordinal["quantity"] = NULL
criconematidae_ordinal["risk_category"] = NULL

x_ordinal_criconematidae = criconematidae_ordinal
x_ordinal_criconematidae = as.matrix(x_ordinal_criconematidae)
class(x_ordinal_criconematidae)

criconematidae_cv = ordinalNetCV(x_ordinal_criconematidae, risk_criconematidae, 
                                  tuneMethod="cvMisclass",family = "cumulative", link = "logit",
                                  lambdaMinRatio = 1e-04,printProgress = FALSE)
summary(criconematidae_cv)
colMeans(summary(criconematidae_cv))

```

### Odd ratio

```{r}


#Soil 2 x 6

criconematidae_log01_ordinal<-crico_predicted$odd[crico_predicted$soil==2]
criconematidae_log02_ordinal<-crico_predicted$odd[crico_predicted$soil==6]

criconematidae_log01_ordinal = mean(criconematidae_log01_ordinal)
criconematidae_log02_ordinal = mean(criconematidae_log02_ordinal)


odd_criconematidae_ordinal= c(criconematidae_log01_ordinal,criconematidae_log02_ordinal)

odd_criconematidae_ordinal = as.data.frame(odd_criconematidae_ordinal)
odd_criconematidae_ordinal$nematode = c("Criconematidae","Criconematidae")
odd_criconematidae_ordinal$comparation = c("2x6","2x6")
colnames(odd_criconematidae_ordinal) = c("odd","nematode","comparation")

soil02_criconematidae = odd_criconematidae_ordinal



#Soil 5x6

criconematidae_log01_ordinal<-crico_predicted$odd[crico_predicted$soil==5]
criconematidae_log02_ordinal<-crico_predicted$odd[crico_predicted$soil==6]

criconematidae_log01_ordinal = mean(criconematidae_log01_ordinal)
criconematidae_log02_ordinal = mean(criconematidae_log02_ordinal)


odd_criconematidae_ordinal= c(criconematidae_log01_ordinal,criconematidae_log02_ordinal)

odd_criconematidae_ordinal = as.data.frame(odd_criconematidae_ordinal)
odd_criconematidae_ordinal$nematode = c("Criconematidae","Criconematidae")
odd_criconematidae_ordinal$comparation = c("5x6","5x6")
colnames(odd_criconematidae_ordinal) = c("odd","nematode","comparation")

soil05_criconematidae = odd_criconematidae_ordinal

#Tillage 1x2

criconematidae_log01_ordinal<-crico_predicted$odd[crico_predicted$tillage==1]
criconematidae_log02_ordinal<-crico_predicted$odd[crico_predicted$tillage==2]

criconematidae_log01_ordinal = mean(criconematidae_log01_ordinal)
criconematidae_log02_ordinal = mean(criconematidae_log02_ordinal)


odd_criconematidae_ordinal= c(criconematidae_log01_ordinal,criconematidae_log02_ordinal)

odd_criconematidae_ordinal = as.data.frame(odd_criconematidae_ordinal)
odd_criconematidae_ordinal$nematode = c("Criconematidae","Criconematidae")
odd_criconematidae_ordinal$comparation = c("tillage","tillage")
colnames(odd_criconematidae_ordinal) = c("odd","nematode","comparation")

tillage_criconematidae = odd_criconematidae_ordinal

#Cropping 1x2

criconematidae_log01_ordinal<-crico_predicted$odd[crico_predicted$cropping==1]
criconematidae_log02_ordinal<-crico_predicted$odd[crico_predicted$cropping==2]

criconematidae_log01_ordinal = mean(criconematidae_log01_ordinal)
criconematidae_log02_ordinal = mean(criconematidae_log02_ordinal)


odd_criconematidae_ordinal= c(criconematidae_log01_ordinal,criconematidae_log02_ordinal)

odd_criconematidae_ordinal = as.data.frame(odd_criconematidae_ordinal)
odd_criconematidae_ordinal$nematode = c("Criconematidae","Criconematidae")
odd_criconematidae_ordinal$comparation = c("cropping","cropping")
colnames(odd_criconematidae_ordinal) = c("odd","nematode","comparation")

cropping_criconematidae = odd_criconematidae_ordinal

all_odd_criconematidae = rbind(soil02_criconematidae,soil05_criconematidae,
                               cropping_criconematidae,
                       tillage_criconematidae)

all_odd_ratio_criconematidae = all_odd_criconematidae %>% 
  group_by(comparation) %>% 
    summarise(
    n = n(),
    ratio_exp = exp(odd[1]/odd[2]),
    ratio = odd[1]/odd[2],
    mean = mean(odd),
    sd = sd(odd),
   se=sd/sqrt(n)) %>% 
  group_by(comparation) %>% 
  mutate(
   lower = exp(ratio - 1.96 * se),
   upper = exp(ratio + 1.96 * se))

coef(summary(crico_ordinal))
```

## Selecting of variables (LASSO) - CONFERIR

```{r,eval = FALSE, echo = TRUE}
library(pROC)
library(ordinalNet)

all_genus = rbind(helicotylenchus_ordinal,pratylenchus_ordinal,criconematidae_ordinal)
all_genus$risk_category = as.numeric(all_genus$risk_category)

all_genus["samples"] = NULL
all_genus = as.data.frame(all_genus)
all_genus["quantity"] = NULL
class(all_genus)

set.seed(123)
lambdas <- 10^seq(2, -3, by = -0.1)


y <- all_genus$risk_category
y = as.factor(y)


X <- as.matrix(subset(all_genus, select = -risk_category))


# Ajuste do modelo LASSO para regressão ordinal
lasso_model <- ordinalNet(X, y,
  alpha = 1,
  lambdaVals = NULL,
  nLambda = 20,
  family = "cumulative",
  standardize = TRUE
)

# Coeficientes do modelo
coef(lasso_model)
summary(lasso_model)

# Extrair os valores de lambda e os erros de validação cruzada

lambda_values <- lasso_model$lambda #0.0004777567

#Best model

lasso_model_best = ordinalNet(X, y,
  alpha = 1,
  lambdaVals = 0.0004777567,
  family = "cumulative",
  standardize = TRUE
)
coef(lasso_model_best)

helic_ordinal = clm(risk_category ~ tillage + cropping + soil + sand + clay  +
                       ph,
                     data = helicotylenchus_ordinal,link = "logit")
car::vif(helic_ordinal)

```

## Distribution

```{r,eval = FALSE, echo = TRUE}
helicotylenchus_threshold = helicotylenchus_ordinal %>% 
  dplyr::group_by(risk_category) %>% 
  dplyr::summarise(
    n = n()
  ) %>% 
  dplyr::group_by(risk_category)%>% 
  dplyr::mutate(
    porcentagem = (n/320)*100
  )

helicotylenchus_ordinal %>% 
  dplyr::filter(risk_category == "2")

heli_threshold = helicotylenchus_threshold %>% 
  ggplot(aes(risk_category,n, fill = risk_category))+
  geom_bar(stat = "identity")+
  scale_fill_viridis_d()+
  theme_minimal()+
  labs(x = "Risk category",
       y = "Samples")+
  theme(text = element_text(size = 16),
        legend.position = "none")



pratylenchus_threshold = pratylenchus_ordinal %>% 
  dplyr::group_by(risk_category) %>% 
 dplyr::summarise(
    n = n()
  ) %>% 
  dplyr::group_by(risk_category) %>% 
  dplyr::mutate(
    porcentagem = (n/320)*100
  )

pratylenchus_ordinal%>% 
  dplyr::filter(risk_category == "2")


praty_threshold = pratylenchus_threshold %>% 
  ggplot(aes(risk_category,n, fill = risk_category))+
  geom_bar(stat = "identity")+
  scale_fill_viridis_d()+
  theme_minimal()+
  labs(x = "Risk category",
       y = "")+
  theme(text = element_text(size = 16),
        legend.position = "none")
  
criconematidae_threshold = criconematidae_ordinal %>% 
  dplyr::group_by(risk_category) %>% 
 dplyr::summarise(
    n = n()
  ) %>% 
  dplyr::group_by(risk_category) %>% 
  dplyr::mutate(
    porcentagem = (n/320)*100
  )

pratylenchus_ordinal%>% 
 dplyr::filter(risk_category == "2")

crico_threshold = criconematidae_threshold %>% 
  ggplot(aes(risk_category,n, fill = risk_category))+
  geom_bar(stat = "identity")+
  scale_fill_viridis_d()+
  theme_minimal()+
  labs(x = "Risk category",
       y = "")+
  theme(text = element_text(size = 16),
        legend.position = "none")

  
  plot_grid(heli_threshold, praty_threshold,crico_threshold,
            labels = c("A","B","C"), label_size = 18, ncol= 3)
  
  ggsave("threshold_nematodes.png", bg = "white", width = 14, height = 8, dpi = 600)
```

## Overall probability

```{r}

helic_predicted$genus = "Helicotylenchus"
praty_predicted$genus = "Pratylenchus"
crico_predicted$genus = "Criconematidae"

nematodes_ordinal = rbind(helic_predicted, praty_predicted, crico_predicted)


nematodes_ordinal_altitude <- nematodes_ordinal %>%
  dplyr::mutate(altitute = case_when(altitute >= 0 ~ altitute,
                                altitute == NA_character_ ~ 0 )) %>% 
  dplyr::mutate(altitude_range = case_when(
    altitute <= 400 ~ "280 a 400",
    altitute <= 600  ~ "401 a 600",
    altitute <= 800 ~ "601 a 800",
    altitute <= 1000 ~ "801 a 1000",
    altitute > 1000 ~"acima de 1001"
  ))

altitude_ordinal = nematodes_ordinal_altitude %>% 
    group_by(genus,altitude_range) %>% 
  summarise(
    
    prob = prob
   
  ) 


point = altitude_ordinal %>% 
     group_by(genus,altitude_range) %>% 
  summarise(
    probi = prob) %>% 
  group_by(genus,altitude_range) %>%
  summarise(
   prob = mean(probi),
   sd = sd(probi))
 


point %>% 
  ggplot(aes(altitude_range,prob))+
  geom_jitter(data = altitude_ordinal, alpha = .7, size = 5, width = .03, color = "gray")+
  #geom_errorbar(aes(altitude_range,ymin = prob - sd , max = prob + sd))+
  #geom_violin(data = altitude_genus, alpha = .7, width = .03, fill = "grey", color = "grey")+
  geom_pointrange(aes(ymin = prob- sd, ymax = prob + sd), color = "black", size = 0.5)+
  theme_minimal()+
  facet_wrap(~genus)+
  labs(x= "Range altitude",
       y = "Probability")+
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 45, hjust = 1))
   
  
```

## Plotting

```{r}
helic_predicted$genus= replicate(320,"Helicotylenchus")
crico_predicted$genus= replicate(320,"Criconematidae")
praty_predicted$genus= replicate(320,"Pratylenchus")


genus_ordinal = rbind(helic_predicted, praty_predicted, crico_predicted)


genus_ordinal[genus_ordinal$risk_category== "0",c("risk")] <- "Not risk"
genus_ordinal[genus_ordinal$risk_category== "1",c("risk")] <- "Insignificant risk"
genus_ordinal[genus_ordinal$risk_category== "2",c("risk")] <- "Moderate-high risk"
```

### By sand

```{r}
sand_ordinal = genus_ordinal %>% 
  filter(risk == "Moderate-high risk") %>% 
  #filter(genus == "Rotylenchulus") %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "grey", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 3)+
  theme_minimal()+
  labs(x = "Sand",
       y = "")+
  theme(text = element_text(size = 18))

#ggsave("probability_ordinal_sand.png", dpi = 600, width = 12, height = 6,  bg = "white")
```

### By clay

```{r}
clay_ordinal = genus_ordinal %>% 
  filter(risk == "Moderate-high risk") %>%
  ggplot(aes(clay,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F,color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 3)+
  theme_minimal()+
  labs(x = "Clay",
       y = "Probability")+
  theme(text = element_text(size = 18))

#ggsave("probability_ordinal_clay.png", dpi = 600, width = 12, height = 6,  bg = "white")
```

### By pH

```{r}
ph_ordinal = genus_ordinal %>% 
  filter(risk == "Moderate-high risk") %>%
  ggplot(aes(ph,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
  facet_wrap(~genus, nrow = 3)+
  theme_minimal()+
  labs(x = "pH",
       y = "")+
  theme(text = element_text(size = 18))

#ggsave("probability_ordinal_ph.png", dpi = 600, width = 12, height = 6,  bg = "white")
```

### Joining

```{r}
(clay_ordinal|sand_ordinal|ph_ordinal)

ggsave("ordinal_genus.png", dpi = 600, bg = "white",height = 6, width = 12)
```

