---
title: "lixo"
format: html
editor: visual
---

```{r}


#incidence = incidence %>% 
 # filter(!city %in% c("DFL","DFC"))


 incidence[incidence$city== "Abadiânia",c("samples_n")] <- 7
 incidence[incidence$city== "Acreuna",c("samples_n")] <- 15
 incidence[incidence$city== "Alexânia",c("samples_n")] <- 10
 incidence[incidence$city== "Alto Paraiso",c("samples_n")] <- 5
 incidence[incidence$city== "Barro Alto",c("samples_n")] <- 8
 incidence[incidence$city== "Bom Jesus de Goiás",c("samples_n")] <- 15
 incidence[incidence$city== "Cabeceiras",c("samples_n")] <- 5
 incidence[incidence$city== "Caiâponia",c("samples_n")] <- 15
 incidence[incidence$city== "Ceres",c("samples_n")] <- 15
 incidence[incidence$city== "Cristalina",c("samples_n")] <- 15
 incidence[incidence$city== "Doverlândia",c("samples_n")] <- 15
 incidence[incidence$city== "Firminópolis",c("samples_n")] <- 5
 incidence[incidence$city== "Formosa",c("samples_n")] <- 5
 incidence[incidence$city== "Jataí",c("samples_n")] <- 15
 incidence[incidence$city== "Luziânia",c("samples_n")] <- 15
 incidence[incidence$city== "Mineiros",c("samples_n")] <- 15
 incidence[incidence$city== "Montividiu",c("samples_n")] <- 15
 incidence[incidence$city== "Nova Crixás",c("samples_n")] <- 5 
 incidence[incidence$city== "Planaltina",c("samples_n")] <- 15
 incidence[incidence$city== "Porangatu",c("samples_n")] <-  15
 incidence[incidence$city== "Portelândia",c("samples_n")] <- 15
 incidence[incidence$city== "Rio Verde",c("samples_n")] <- 15
 incidence[incidence$city== "São Miguel do Passo Quatro",c("samples_n")] <- 15
 incidence[incidence$city== "Serranópolis",c("samples_n")] <- 10
 incidence[incidence$city== "Silvânia",c("samples_n")] <- 15
 incidence[incidence$city== "Vianópolis",c("samples_n")] <- 15
```


## Models diagnostic (graphic)

```{r}
check_singularity(praty_log)
check_model(praty_log)

simulationOutput <- simulateResiduals(fittedModel = praty_log)
plot(simulationOutput)
```

### Aphelenchus

```{r}

#Model log with best
aphelenchus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchus_log,family = binomial(link = "logit"))

summary(aphelenchus_binomial)

pred_aphelenchus <- predict(aphelenchus_binomial, aphelenchus_log, type = "response")


#Better thresholds
dat_aphelenchus <- data.frame(1, aphelenchus_log$incidence, pred_aphelenchus)
  optimal.thresholds(dat_aphelenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_aphelenchus > 0.82)),  mode= "everything",  reference = as.factor(aphelenchus_log$incidence))


#Plot
auc.roc.plot(dat_aphelenchus)
calibration.plot(dat_aphelenchus)

error.threshold.plot(dat_aphelenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_aphelenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_aphelenchus,threshold=0.82,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_aphelenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_aphelenchus <- roc.plot.calculate(dat_aphelenchus) #data frame

roc_aphelenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_aphelenchus <- NULL
for(i in 1:nrow(aphelenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- aphelenchus_log[-i,]
    test <- aphelenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.82,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_aphelenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_aphelenchus)


```

### Aphelenchoides

```{r}
#Model log with best
aphelenchoides_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchoides_log,family = binomial(link = "logit"))

summary(aphelenchoides_binomial)

pred_aphelenchoides <- predict(aphelenchoides_binomial, aphelenchoides_log, type = "response")

#Better thresholds
dat_aphelenchoides <- data.frame(1, aphelenchoides_log$incidence, pred_aphelenchoides)
optimal.thresholds(dat_aphelenchoides)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_aphelenchoides > 0.65)),  mode= "everything",  reference = as.factor(aphelenchoides_log$incidence))


#Plot
auc.roc.plot(dat_aphelenchoides)
calibration.plot(dat_aphelenchoides)

error.threshold.plot(dat_aphelenchoides, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_aphelenchoides), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_aphelenchoides,threshold=0.65,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_aphelenchoides) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_aphelenchoides <- roc.plot.calculate(dat_aphelenchoides) #data frame

roc_aphelenchoides |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_aphelenchoides <- NULL
for(i in 1:nrow(aphelenchoides_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- aphelenchoides_log[-i,]
    test <- aphelenchoides_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = aphelenchoides_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.65,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_aphelenchoides[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_aphelenchoides)


```

### Criconematidae

```{r}
#Model log with best
criconematidae_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = criconematidae_log,family = binomial(link = "logit"))

summary(criconematidae_binomial)

pred_criconematidae <- predict(criconematidae_binomial, criconematidae_log, type = "response")

#Better thresholds
dat_criconematidae <- data.frame(1, criconematidae_log$incidence, pred_criconematidae)
optimal.thresholds(dat_criconematidae)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_criconematidae > 0.26)),  mode= "everything",  reference = as.factor(criconematidae_log$incidence))


#Plot
auc.roc.plot(dat_criconematidae)
calibration.plot(dat_criconematidae)

error.threshold.plot(dat_criconematidae, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_criconematidae), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_criconematidae,threshold=0.26,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_criconematidae) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_criconematidae <- roc.plot.calculate(dat_criconematidae) #data frame

roc_criconematidae |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_criconematidae <- NULL
for(i in 1:nrow(criconematidae_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- criconematidae_log[-i,]
    test <- criconematidae_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = criconematidae_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.26,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_criconematidae[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_criconematidae)


```

### Helicotylenchus

```{r}
#Model log with best
helicotylenchus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = helicotylenchus_log,family = binomial(link = "logit"))

summary(helicotylenchus_binomial)

pred_helicotylenchus <- predict(helicotylenchus_binomial, helicotylenchus_log, type = "response")

#Better thresholds
dat_helicotylenchus <- data.frame(1, helicotylenchus_log$incidence, pred_helicotylenchus)
optimal.thresholds(dat_helicotylenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_helicotylenchus > 0.9)),  mode= "everything",  reference = as.factor(helicotylenchus_log$incidence))


#Plot
auc.roc.plot(dat_helicotylenchus)
calibration.plot(dat_helicotylenchus)

error.threshold.plot(dat_helicotylenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_helicotylenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_helicotylenchus,threshold=.9,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_helicotylenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_helicotylenchus <- roc.plot.calculate(dat_helicotylenchus) #data frame

roc_helicotylenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_helicotylenchus <- NULL
for(i in 1:nrow(helicotylenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- helicotylenchus_log[-i,]
    test <- helicotylenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = helicotylenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.9,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_helicotylenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_helicotylenchus)



```

### Rotylenchulus

```{r}
#Model log with best
rotylenchulus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = rotylenchulus_log,family = binomial(link = "logit"))

summary(rotylenchulus_binomial)

pred_rotylenchulus <- predict(rotylenchulus_binomial, rotylenchulus_log, type = "response")

#Better thresholds
dat_rotylenchulus <- data.frame(1, rotylenchulus_log$incidence, pred_rotylenchulus)
optimal.thresholds(dat_rotylenchulus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_rotylenchulus > .06)),  mode= "everything",  reference = as.factor(rotylenchulus_log$incidence))


#Plot
auc.roc.plot(dat_rotylenchulus)
calibration.plot(dat_rotylenchulus)

error.threshold.plot(dat_rotylenchulus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_rotylenchulus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_rotylenchulus,threshold=.06,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_rotylenchulus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_rotylenchulus <- roc.plot.calculate(dat_rotylenchulus) #data frame

roc_rotylenchulus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_rotylenchulus <- NULL
for(i in 1:nrow(rotylenchulus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- rotylenchulus_log[-i,]
    test <- rotylenchulus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = rotylenchulus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > .06,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_rotylenchulus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_rotylenchulus)

```

### Pratylenchus

```{r}
#Model log with best
pratylenchus_binomial <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = pratylenchus_log,family = binomial(link = "logit"))

summary(pratylenchus_binomial)

pred_pratylenchus <- predict(pratylenchus_binomial, pratylenchus_log, type = "response")

#Better thresholds
dat_pratylenchus <- data.frame(1, pratylenchus_log$incidence, pred_pratylenchus)
optimal.thresholds(dat_pratylenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_pratylenchus > 0.58)),  mode= "everything",  reference = as.factor(pratylenchus_log$incidence))


#Plot
auc.roc.plot(dat_pratylenchus)
calibration.plot(dat_pratylenchus)

error.threshold.plot(dat_pratylenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_pratylenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_pratylenchus,threshold=.58,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_pratylenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_pratylenchus <- roc.plot.calculate(dat_pratylenchus) #data frame

roc_pratylenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_pratylenchus <- NULL
for(i in 1:nrow(pratylenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- pratylenchus_log[-i,]
    test <- pratylenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = pratylenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.58,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_pratylenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_pratylenchus)



```

## Joining

```{r}

loocv_all = cbind(acc_aphelenchoides,acc_aphelenchus,acc_criconematidae,
                  acc_helicotylenchus,acc_pratylenchus,acc_rotylenchulus)
```

## Altitude (REMOVER)

```{r}

helicotylenchus_log$prob = predict(heli_log, type = "response")
criconematidae_log$prob = predict(crico_log, type = "response")
aphelenchus_log$prob = predict(aphel_log, type = "response")
aphelenchoides_log$prob = predict(aphels_log, type = "response")
pratylenchus_log$prob = predict(praty_log, type = "response")
rotylenchulus_log$prob = predict(rotyl_log, type = "response")
#exp(summary(heli_log))


altitude = rbind(helicotylenchus_log,criconematidae_log,aphelenchus_log,aphelenchoides_log,
                 pratylenchus_log,rotylenchulus_log)


altitude1 <- altitude %>%
  mutate(altitute = case_when(altitute >= 0 ~ altitute,
                                altitute == NA_character_ ~ 0 )) %>% 
  mutate(altitude_range = case_when(
    altitute <= 400 ~ "280 to 400",
    altitute <= 600  ~ "401 to 600",
    altitute <= 800 ~ "601 to 800",
    altitute <= 1000 ~ "801 to 1000",
    altitute > 1000 ~"Above 1001"
  ))

altitude_genus = altitude1 %>% 
    group_by(genus,altitude_range) %>% 
  summarise(
    
    prob = prob)
   


point = altitude1 %>% 
    group_by(genus,altitude_range) %>% 
  summarise(
    
    probi = mean(prob),
    sd = sd(prob)
   
  ) %>%
  group_by(genus,altitude_range) %>% 
  summarise(
   prob = mean(probi),
   sd = sd
   )
 
#write_xlsx(point,"THAIS/point.xlsx")


point %>% 
    ggplot(aes(altitude_range,prob))+
  geom_jitter(data = altitude_genus, alpha = .2, size = 5, width = .03, color = "gray")+
  #geom_violin(data = altitude_genus, alpha = .7, width = .03, fill = "grey", color = "grey")+
  geom_pointrange(aes(ymin = prob- sd, ymax = prob + sd), color = "#041e39", size = 1)+
  ggthemes::theme_few()+
  facet_wrap(~genus)+
  labs(x= "Altitude (m)",
       y = "Probability")+
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(face = "bold"),
        strip.text = element_text(face = "italic"))
  
#ggsave("fig/prob_altitude.png", bg = "white", width = 12, height = 8)
```

## Overall

```{r}
altitude_genus = altitude1 %>% 
    group_by(genus,altitude_range,sand,clay,ph,cropping,soil,tillage) %>% 
  summarise(
    
    prob = prob
   
  )

#write_xlsx(altitude_genus,"THAIS/altitude_genus.xlsx")

```

## Probability of incidence

### By tillage and sand

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+tillage, nrow = 3)+
  ggthemes::theme_few()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_sand_tillage_all_genus.png", dpi = 600, width = 14, height = 8, bg = "white")
```

### By tillage and  pH

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(ph,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+tillage, nrow = 3)+
  theme_minimal()+
  labs(x = "pH gradient",
       y = "Probabilitiy")+
  theme(text = element_text(size = 18))

ggsave("probability_ph_tillage_all_genus.png", dpi = 600, width = 14, height = 8,bg = "white")
```

### By tillage and clay

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(clay,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+tillage, nrow = 3)+
  theme_minimal()+
  labs(x = "Clay",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_clay_tillage_all_genus.png", dpi = 600, width = 14, height = 8, bg = "white")
```

### By sand and soil type

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F,color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+soil, nrow = 3)+
  theme_minimal()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_soil_all_genus.png", dpi = 600, width = 14,height = 8, bg = "white")
```

### By cropping and sand

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+cropping, nrow = 3)+
  theme_minimal()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_cropping_sand_all_genus.png", dpi = 600, width = 14, height = 8, bg = "white")
```

### By cropping and pH

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(ph,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black",fullrange = T)+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus+cropping, nrow = 3)+
  theme_minimal()+
  labs(x = "pH gradient",
       y = "Probability")+
  theme(text = element_text(size = 18))

ggsave("probability_cropping_ph_all_genus.png", dpi = 600, width = 14,height = 6, bg = "white")
```

### By sand

```{r}
altitude_genus %>% 
  #filter(!genus %in% c("Aphelenchus","Criconematidae","Rotylenchulus")) %>% 
  ggplot(aes(sand,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 2)+
  ggthemes::theme_few()+
  labs(x = "Sand",
       y = "Probability")+
  theme(text = element_text(size = 18),
        axis.title = element_text(face = "bold"))

ggsave("fig/probability_sand_genus.png", dpi = 600, width = 14, height = 6, bg = "white")
```

### By pH

```{r}
altitude_genus %>% 
  #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(ph,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 2)+
  ggthemes::theme_few()+
  labs(x = "pH",
       y = "Probability")+
  theme(text = element_text(size = 18),
        axis.title = element_text(face = "bold"))

ggsave("fig/probability_ph_genus.png", dpi = 600, width = 14,height = 6, bg = "white")
```

### By clay

```{r}
altitude_genus %>% 
 #filter(genus == "Aphelenchus") %>% 
  ggplot(aes(clay,prob))+
  geom_jitter(alpha = .9, size = 3, color = "gray", width = 0.01, height = 0.001)+
 geom_smooth(method = "glm", size = 2, se = F, color = "black")+
 # geom_line(data = gg2, aes(silt = x, y = ymin), size = 0.02) +
  #  geom_line(data = gg2, aes(silt = x, y = ymax), size = 0.02)+
  facet_wrap(~genus, nrow = 2)+
  ggthemes::theme_few()+
  labs(x = "Clay",
       y = "Probability")+
  theme(text = element_text(size = 18),
        axis.title = element_text(face = "bold"))

ggsave("fig/probability_clay_genus.png", dpi = 600, width = 14,height = 6, bg = "white")
```

# Random efect - Incidence

## Estimation

```{r}
pratylenchus_groups_binomial = ranef(pratylenchus_glmer_binomial)
pratylenchus_groups_binomial = as.data.frame(pratylenchus_groups_binomial)

helicotylenchus_groups_binomial = ranef(helicotylenchus_glmer_binomial)
helicotylenchus_groups_binomial = as.data.frame(helicotylenchus_groups_binomial)

rotylenchulus_groups_binomial = ranef(rotylenchulus_glmer_binomial)
rotylenchulus_groups_binomial = as.data.frame(rotylenchulus_groups_binomial)

criconematidae_groups_binomial = ranef(criconematidae_glmer_binomial)
criconematidae_groups_binomial = as.data.frame(criconematidae_groups_binomial)

aphelenchus_groups_binomial = ranef(aphelenchus_glmer_binomial)
aphelenchus_groups_binomial = as.data.frame(aphelenchus_groups_binomial)

aphelenchoides_groups_binomial = ranef(aphelenchoides_glmer_binomial)
aphelenchoides_groups_binomial = as.data.frame(aphelenchoides_groups_binomial)





all_genus_binomial = rbind(pratylenchus_groups_binomial,helicotylenchus_groups_binomial,
                  rotylenchulus_groups_binomial,criconematidae_groups_binomial,
                  aphelenchus_groups_binomial,aphelenchoides_groups_binomial)

nrow(aphelenchoides_groups_binomial)

pratylenchus_name_binomial = replicate(21,"Pratylenchus")
pratylenchus_name_binomial = as.data.frame(pratylenchus_name_binomial)
colnames(pratylenchus_name_binomial)[1] = "genus"

helicotylenchus_name_binomial = replicate(21,"Helicotylenchus")
helicotylenchus_name_binomial = as.data.frame(helicotylenchus_name_binomial)
colnames(helicotylenchus_name_binomial)[1] = "genus"

rotylenchulus_name_binomial = replicate(21,"Rotylenchulus")
rotylenchulus_name_binomial = as.data.frame(rotylenchulus_name_binomial)
colnames(rotylenchulus_name_binomial)[1] = "genus"

criconematidae_name_binomial = replicate(21,"Criconematidae")
criconematidae_name_binomial = as.data.frame(criconematidae_name_binomial)
colnames(criconematidae_name_binomial)[1] = "genus"

aphelenchus_name_binomial = replicate(21,"Aphelenchus")
aphelenchus_name_binomial = as.data.frame(aphelenchus_name_binomial)
colnames(aphelenchus_name_binomial)[1] = "genus"

aphelenchoides_name_binomial = replicate(21,"Aphelenchoides")
aphelenchoides_name_binomial = as.data.frame(aphelenchoides_name_binomial)
colnames(aphelenchoides_name_binomial)[1] = "genus"




genus_binomial = rbind(pratylenchus_name_binomial,helicotylenchus_name_binomial,
                    rotylenchulus_name_binomial,criconematidae_name_binomial,
                    aphelenchus_name_binomial,aphelenchoides_name_binomial)

all_genus_binomial = cbind(all_genus_binomial,genus_binomial)

```

## Plotting

```{r}


select_region_city_binomial = all_genus_binomial %>% 
    filter(genus %in% c("Helicotylenchus","Pratylenchus","Aphelenchus",
                      "Aphelenchoides","Rotylenchulus", "Criconematidae")) %>% 
  filter(grpvar == "region:city")


all_genus_binomial1 <- all_genus_binomial %>% #CORRIGIR O P-VALOR E OBTER PELO MÉTODO DE PSEUDO-VEROSSIMILHANÇA 
  group_by(genus) %>%
  summarise( 
    n=n(),
    mean=mean(condval),
    sd=sd(condval)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.05)/2 + .5, n-1),
          lower = (mean - 1.96*se),
          upper = (mean + 1.96*se),
          p_value = ((sd/sqrt(n)) * 1.96))


summary(pratylenchus_glmer_binomial)
confint(helicotylenchus_glmer_binomial)


incidence = all_genus_binomial1 |> 
  #filter(!genus == "Pratylenchus") |> 
  ggplot(aes(reorder(genus,+mean), y = mean))+
  geom_bar(stat = "identity", position = "dodge", fill = "darkgreen")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1,vjust = 1.2, size = 14),
        text = element_text(size = 16))+
  labs(x = "Morphological group",
       y = "Random effect",
       title = "Municipality/Region")

```

# Abundance - ordinal

### Cross-validation

```{r}
library(ordinalNet)

risk_helicotylenchus = helicotylenchus_ordinal$risk_category
class(risk_helicotylenchus)

helicotylenchus_ordinal_cv = helicotylenchus_ordinal

helicotylenchus_ordinal_cv["samples"] = NULL
helicotylenchus_ordinal_cv["ec"] = NULL
helicotylenchus_ordinal_cv["altitute"] = NULL
helicotylenchus_ordinal_cv["quantity"] = NULL
helicotylenchus_ordinal_cv["risk_category"] = NULL

x_ordinal_helicotylenchus = helicotylenchus_ordinal_cv
x_ordinal_helicotylenchus = as.matrix(x_ordinal_helicotylenchus)
class(x_ordinal_helicotylenchus)

helicotylenchus_cv = ordinalNetCV(x_ordinal_helicotylenchus, risk_helicotylenchus,
                                  tuneMethod=c("cvMisclass"),family = "cumulative", link = "logit",
                                  lambdaMinRatio = 1e-04,printProgress = FALSE)
summary(helicotylenchus_cv)
colMeans(summary(helicotylenchus_cv))

```

Helicotylenchus model diagnostic (NÃO ESTÁ AJUSTANDO IGUAL O DE INCIDÊNCIA)

```{r,eval = FALSE, echo = TRUE}
#Model log with best
helic_ordinal = clm(risk_category ~ tillage + cropping + soil + sand + clay  +
                       ph,
                     data = helicotylenchus_ordinal,link = "logit")


summary(helic_ordinal)

helic_predicted1 = helic_predicted |> 
  filter(!risk_category == 2)

pred_helicotylenchus_ordinal <- helic_predicted1$prob


  


#Better thresholds
dat_helicotylenchus <- data.frame(1,helic_predicted1$risk_category, pred_helicotylenchus_ordinal)
dat_helicotylenchus[dat_helicotylenchus$helic_predicted1.risk_category== "0",c("risk")] <- 0
dat_helicotylenchus[dat_helicotylenchus$helic_predicted1.risk_category== "1",c("risk")] <-1
dat_helicotylenchus["helic_predicted1.risk_category"] = NULL

optimal.thresholds(dat_helicotylenchus)

# Confusion matrix with the MAXsen+spe threshold
confusionMatrix(data = as.factor(as.numeric(pred_helicotylenchus_ordinal > 0.9)),  mode= "everything",  reference = as.factor(helicotylenchus_log$incidence))


#Plot
auc.roc.plot(dat_helicotylenchus)
calibration.plot(dat_helicotylenchus)

error.threshold.plot(dat_helicotylenchus, opt.methods = 3) #takes a single model and plots the sensitivity and specificity as a function of threshold. 
pcc(cmx(dat_helicotylenchus), st.dev = TRUE) #calculates the percent correctly classified and (optionally) the associated standard deviation from a confusion matrix.

presence.absence.accuracy(dat_helicotylenchus,threshold=.9,st.dev=FALSE) #Calculates five accuracy measures (pcc, sensitivity, specificity, Kappa, and AUC)
presence.absence.summary(dat_helicotylenchus) #Produces four types of Presence/Absence accuracy plots for a single set of model Predictions

roc_helicotylenchus <- roc.plot.calculate(dat_helicotylenchus) #data frame

roc_helicotylenchus |> 
  ggplot(aes(threshold, sensitivity))+
  geom_line()+
  geom_line(aes(threshold, specificity))+
  geom_line(aes(threshold, Kappa))


#LOOCV
acc_helicotylenchus <- NULL
for(i in 1:nrow(helicotylenchus_log))
{
    # Train-test splitting
    # 499 samples -> fitting
    # 1 sample -> testing
    train <- helicotylenchus_log[-i,]
    test <- helicotylenchus_log[i,]
    
    # Fitting
    model <- glm(incidence ~ altitute + tillage_corrigido + cropping_corrigido + soil_corrigido + sand + clay + ph, data = helicotylenchus_log,
                family = binomial(link = "logit"))
    
    # Predict results
    results_prob <- predict(model,test,type='response')
    
    # If prob > 0.5 then 1, else 0
    results <- ifelse(results_prob > 0.9,1,0)
    
    # Actual answers
    answers <- test$incidence
    
    # Calculate accuracy
    misClasificError <- mean(answers != results)
    
    # Collecting results
    acc_helicotylenchus[i] <- 1-misClasificError
}

# Average accuracy of the model
mean(acc_helicotylenchus)



```

### Cross-validation

```{r}

risk_pratylenchus = pratylenchus_ordinal$risk_category
class(risk_pratylenchus)

pratylenchus_ordinal_cv = pratylenchus_ordinal

pratylenchus_ordinal_cv["samples"] = NULL
pratylenchus_ordinal_cv["ec"] = NULL
pratylenchus_ordinal_cv["altitute"] = NULL
pratylenchus_ordinal_cv["quantity"] = NULL
pratylenchus_ordinal_cv["risk_category"] = NULL

x_ordinal_pratylenchus = pratylenchus_ordinal_cv
x_ordinal_pratylenchus = as.matrix(x_ordinal_pratylenchus)
class(x_ordinal_pratylenchus)

pratylenchus_cv = ordinalNetCV(x_ordinal_pratylenchus, risk_pratylenchus,family = "cumulative", link = "logit",
                                  tuneMethod=c("cvMisclass"),lambdaMinRatio = 1e-04,printProgress = FALSE)
summary(pratylenchus_cv)
colMeans(summary(pratylenchus_cv))

```

### Cross-validation

```{r}

risk_criconematidae = criconematidae_ordinal$risk_category
class(risk_criconematidae)

criconematidae_ordinal["samples"] = NULL
criconematidae_ordinal["ec"] = NULL
criconematidae_ordinal["altitute"] = NULL
criconematidae_ordinal["quantity"] = NULL
criconematidae_ordinal["risk_category"] = NULL

x_ordinal_criconematidae = criconematidae_ordinal
x_ordinal_criconematidae = as.matrix(x_ordinal_criconematidae)
class(x_ordinal_criconematidae)

criconematidae_cv = ordinalNetCV(x_ordinal_criconematidae, risk_criconematidae, 
                                  tuneMethod="cvMisclass",family = "cumulative", link = "logit",
                                  lambdaMinRatio = 1e-04,printProgress = FALSE)
summary(criconematidae_cv)
colMeans(summary(criconematidae_cv))

```


# Modelagem da abundância

```{r}
summary(aphelenchus_glmer)
check_singularity(aphelenchus_glmer)
check_overdispersion(aphelenchus_glmer)
model_performance(aphelenchus_glmer)
check_model(aphelenchus_glmer)

#ggsave("model_check/check_aphelenchus_abundance_glmm.png", bg = "white", dpi = 600, width = 12)



aphelenchus_res <- simulateResiduals(aphelenchus_glmer)
plot(aphelenchus_res)




###################

summary(aphelenchoides_glmer)
check_singularity(aphelenchoides_glmer)
check_overdispersion(aphelenchoides_glmer)
model_performance(aphelenchoides_glmer)
check_model(aphelenchoides_glmer)

#ggsave("model_check/check_aphelenchoides_abundance_glmm.png", bg = "white", dpi = 600, width = 12)


aphelenchoides_res <- simulateResiduals(aphelenchoides_glmer)
plot(aphelenchoides_res)

####################


summary(pratylenchus_glmer)
check_singularity(pratylenchus_glmer)
check_overdispersion(pratylenchus_glmer)
model_performance(pratylenchus_glmer)
check_model(pratylenchus_glmer)

#ggsave("model_check/check_pratylenchus_abundance_glmm.png", bg = "white", dpi = 600, width = 12)


pratylenchus_res <- simulateResiduals(pratylenchus_glmer)
plot(pratylenchus_res)

######################


summary(helicotylenchus_glmer)
check_singularity(helicotylenchus_glmer)
check_overdispersion(helicotylenchus_glmer)
model_performance(helicotylenchus_glmer)
check_model(helicotylenchus_glmer)

#ggsave("model_check/check_helicotylenchus_abundance_glmm.png", bg = "white", dpi = 600, width = 12)


helicotylenchus_res <- simulateResiduals(helicotylenchus_glmer)
plot(helicotylenchus_res)


#######################



summary(rotylenchulus_glmer)
check_singularity(rotylenchulus_glmer)
check_overdispersion(rotylenchulus_glmer)
model_performance(rotylenchulus_glmer)
check_model(rotylenchulus_glmer)

#ggsave("model_check/check_rotylenchulus_abundance_glmm.png", bg = "white", dpi = 600, width = 12)


rotylenchulus_res <- simulateResiduals(rotylenchulus_glmer)
plot(rotylenchulus_res)

######################



summary(criconematidae_glmer)
check_singularity(criconematidae_glmer)
check_overdispersion(criconematidae_glmer)
model_performance(criconematidae_glmer)
check_model(criconematidae_glmer)

#ggsave("model_check/check_rotylenchulus_abundance_glmm.png", bg = "white", dpi = 600, width = 12)


criconematidae_res <- simulateResiduals(criconematidae_glmer)
plot(criconematidae_res)
```

# Random efect - Abundance

## Estimation

```{r}
pratylenchus_groups = ranef(pratylenchus_glmer)
pratylenchus_groups = as.data.frame(pratylenchus_groups)

helicotylenchus_groups = ranef(helicotylenchus_glmer)
helicotylenchus_groups = as.data.frame(helicotylenchus_groups)

rotylenchulus_groups = ranef(rotylenchulus_glmer)
rotylenchulus_groups = as.data.frame(rotylenchulus_groups)

criconematidae_groups = ranef(criconematidae_glmer)
criconematidae_groups = as.data.frame(criconematidae_groups)

aphelenchus_groups = ranef(aphelenchus_glmer)
aphelenchus_groups = as.data.frame(aphelenchus_groups)

aphelenchoides_groups = ranef(aphelenchoides_glmer)
aphelenchoides_groups = as.data.frame(aphelenchoides_groups)





all_genus = rbind(pratylenchus_groups,helicotylenchus_groups,rotylenchulus_groups,
                  criconematidae_groups,aphelenchus_groups,aphelenchoides_groups)

nrow(aphelenchoides_groups)

pratylenchus_name = replicate(217,"Pratylenchus")
pratylenchus_name = as.data.frame(pratylenchus_name)
colnames(pratylenchus_name)[1] = "genus"

helicotylenchus_name = replicate(286,"Helicotylenchus")
helicotylenchus_name = as.data.frame(helicotylenchus_name)
colnames(helicotylenchus_name)[1] = "genus"

rotylenchulus_name = replicate(22,"Rotylenchulus")
rotylenchulus_name = as.data.frame(rotylenchulus_name)
colnames(rotylenchulus_name)[1] = "genus"

criconematidae_name = replicate(76,"Criconematidae")
criconematidae_name = as.data.frame(criconematidae_name)
colnames(criconematidae_name)[1] = "genus"

aphelenchus_name = replicate(259,"Aphelenchus")
aphelenchus_name = as.data.frame(aphelenchus_name)
colnames(aphelenchus_name)[1] = "genus"

aphelenchoides_name = replicate(209,"Aphelenchoides")
aphelenchoides_name = as.data.frame(aphelenchoides_name)
colnames(aphelenchoides_name)[1] = "genus"




genus = rbind(pratylenchus_name,helicotylenchus_name,rotylenchulus_name,
              criconematidae_name,aphelenchus_name,aphelenchoides_name)

all_genus = cbind(all_genus,genus)

select_region_city_samples = all_genus %>% 
    filter(genus %in% c("Helicotylenchus","Pratylenchus","Aphelenchus",
                      "Aphelenchoides","Rotylenchulus", "Criconematidae")) %>% 
  filter(grpvar == "region:city:samples")

select_region_city = all_genus %>% 
    filter(genus %in% c("Helicotylenchus","Pratylenchus","Aphelenchus",
                      "Aphelenchoides","Rotylenchulus", "Criconematidae")) %>% 
  filter(grpvar == "region:city")



colnames(select_region_city_samples) = c("group","term","grp","mean","sd","genus")
colnames(select_region_city) = c("group","term","grp","mean","sd","genus")

all_genus1 <- all_genus %>%
  group_by(grpvar,genus) %>%
  summarise( 
    n=n(),
    mean=mean(condval),
    sd=sd(condval)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.05)/2 + .5, n-1),
    lower = (mean - 1.96*se),
    upper = (mean + 1.96*se))
    #p_value = ((sd/sqrt(n)) * 1.96))

```

## Plotting
```{r}
all_genus1 %>% 
    filter(genus %in% c("Helicotylenchus","Pratylenchus","Aphelenchus",
                      "Aphelenchoides","Rotylenchulus", "Criconematidae")) %>% 
 # filter(grpvar == "soil:city:altitute") %>% 
  ggplot(aes(genus,mean))+
  geom_bar(stat = "identity", position = "dodge")+
  facet_wrap(~grpvar)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Environmental

```{r}
summary(praty_log)
check_singularity(praty_log)
model_performance(praty_log)
check_model(praty_log)



check_model(heli_log)
model_performance(heli_log)


check_model(rotyl_log)
model_performance(rotyl_log)

check_model(rotyl_log)
model_performance(rotyl_log)

check_model(aphel_log)
model_performance(aphel_log)


check_model(aphels_log)
model_performance(aphels_log)
```

